### 2023-04-19 1a_Difference Scores paired samples ttest_HowTo.qmd
rstudio test commit 4

---
title: "DiffScoreExample"
---

# Paired Samples T-test {.unnumbered}

## Overview

<p>

This example examines the question whether participants reported
different mean levels of externalizing problems on the CBCL when
measured at Baseline (T0) compared to when measured at the 1-Year
follow-up (T1). This analysis is conducted in two primary steps: 1)
computing a difference score (DiffScore = CBCL EXT T2 - CBCL EXT T1); 2)
conducting a paired samples t-test on the difference score. Our primary
aim is to determine whether average CBCL externalizing scores of the
participants change from assessement 1 (T1) to assessement 2 (T2), while
accounting for observations that are clustered within youth over time.

</p>