### 2023-04-19 2b_Residualized Change Score multiple regression_HowTo.qmd

---
title: "Residualized Change Score Example"
---

# Mulitple Regression
testing testing testing

## Overview

<p>
insert text
</p>