### 2023-04-19 11_Tutorials_MultivariateLatentGrowthCurveModels.qmd
rstudio 9

---
title: "Multivariate Latent Growth Curves"
---