### 2023-04-19 4_Tutorials_SignedRankTest.qmd

---
title: "Signed-Rank Test"
author: "Biostatistics Working Group"
---

## Overview
The longitudinal signed-rank test is a nonparametric alternative to the paired t-test and is appropriate when the data are not normally distributed or when the assumptions of the paired t-test are violated. The test is based on the signed-rank of the differences between the paired observations, and it tests the null hypothesis that the median of the differences is zero. The repeated-measures design allows for the assessment of within-subject changes over time or under different conditions. The signed-rank test is robust to outliers and does not assume a normal distribution of the differences. This method has been shown to have good statistical power and efficiency, particularly when the sample size is small or the distribution is heavily skewed (Erceg-Hurn & Mirosevich, 2008; Bakdash & Marusich, 2017; Garcia-Berthou & Alcaraz, 2004). The signed-rank test is a useful tool for analyzing paired data when the assumption of normality is violated or when the data is highly skewed or contains outliers.

[+add diagrams/figures]

### When to use a Signed-Rank Test?
You should use a Wilcoxon Signed-Rank Test in the following scenario:
1.	You want to know if two groups are different on your variable of interest
2.	Your variable of interest is continuous
3.	You have two and only two groups
4.	You have independent samples
5.	You have a skewed variable of interest

### Getting Started with Longitudinal Signed-Rank Test

In this tutorial, we will introduce the concept of the longitudinal signed-rank test and guide you through a simple example using a small dataset. By the end of this tutorial, you will be able to:

1. Understand the basic concepts of the longitudinal signed-rank test.
2. Perform a longitudinal signed-rank test using example data in R.
3. Interpret the results of the test.

## Basic Example

For this tutorial, we will use a simple dataset containing the following information for a group of individuals:

- Subject ID
- Time point 1 (T1) measurement
- Time point 2 (T2) measurement

| Subject ID | T1  | T2  |
|------------|-----|-----|
| 1          | 100 | 105 |
| 2          | 95  | 100 |
| 3          | 110 | 112 |

### Model Specification and Estimation

We will use the `wilcox.test()` function in R to perform a longitudinal signed-rank test. First, create a data frame with the example data:

```{r}
data <- data.frame(
  subject_id = c(1, 2, 3),
  t1 = c(100, 95, 110),
  t2 = c(105, 100, 112)
)
```

Perform the longitudinal signed-rank test using the wilcox.test() function with the paired = TRUE argument:

```{r}
test_result <- wilcox.test(data$t1, data$t2, paired = TRUE)
```

### Interpreting the Results
To interpret the results of the longitudinal signed-rank test, examine the p-value from the test_result object:

```{r}
test_result$p.value
```

The p-value represents the probability of observing the data if there is no difference between the measurements at T1 and T2. If the p-value is less than a predetermined significance level (e.g., 0.05), we can reject the null hypothesis that there is no difference between the two time points and conclude that there is a significant difference in the measurements between T1 and T2.