### 2023-04-19 5_Tutorials_MarginalModels.qmd

---
title: "Marginal Models"
author: "Biostatistics Working Group"
---

## Overview
Marginal models are a statistical method used to analyze longitudinal or clustered data. The marginal model estimates the average effect of the independent variables on the outcome while accounting for the within-subject correlation, and allows for the estimation of population-averaged effects, in contrast to subject-specific effects estimated by random-effects models (Verbeke & Molenberghs, 2000). The method is similar to the GEE approach but uses a different estimation technique and does not account for subject-specific effects. Marginal models have been shown to be robust to non-normality and non-constant variance, and can handle unbalanced or unequally spaced data (Fitzmaurice et al., 2011). The term marginal in this context is used to emphasize that the model for the mean response at each occasion depends only on the covariates of interest, and not on any random effects or previous responses.

[+add diagrams/figures]

### When to use Longitudinal Marginal Models
You should use xxxxxxx in the following scenario:
1.	You want to know xxxxx
2.	Your variable xxxxx
3.	You have xxxxx

### Getting Started with Longitudinal Marginal Models

In this tutorial, we will introduce the concept of longitudinal marginal models and guide you through a simple example using a small dataset. By the end of this tutorial, you will be able to:

1. Understand the basic concepts of longitudinal marginal models.
2. Fit a longitudinal marginal model using example data in R.
3. Interpret the results of the model.

## Basic Example

For this tutorial, we will use a simple dataset containing the following information for a group of individuals:

- Subject ID
- Time (in years)
- Binary outcome variable (e.g., presence or absence of a particular condition)

| Subject ID | Time | Outcome |
|------------|------|---------|
| 1          | 0    | 0       |
| 1          | 1    | 1       |
| 1          | 2    | 1       |
| 2          | 0    | 0       |
| 2          | 1    | 1       |

### Model Specification and Estimation

We will use the `geepack` package in R to fit a longitudinal marginal model. First, install and load the required package:

```{r}
if (!("lme4" %in% installed.packages())) install.packages("geepack")
library(geepack)
```

Next, create a data frame with the example data:

```{r}
data <- data.frame(
  subject_id = c(1, 1, 1, 2, 2),
  time = c(0, 1, 2, 0, 1),
  outcome = c(0, 1, 1, 0, 1)
)
```

Fit the longitudinal marginal model with a binary outcome using a logit link function:

```{r}
model <- geeglm(outcome ~ time, data = data, id = subject_id, family = binomial(link = "logit"), corstr = "exchangeable")
```

### Interpreting the Results
To interpret the results of the longitudinal marginal model, we can use the summary() function in R to display the estimated population-averaged effects:

```{r}
summary(model)
```

The output will show the estimated effect of time on the binary outcome variable, expressed as an odds ratio. By examining the odds ratio and its associated p-value, we can determine if the change in the outcome variable over time is significant at the population level.