### 2023-04-19 6_Tutorials_GeneralizedEstimatingEquations.qmd

---
title: "Generalized Estimating Equations"
author: "Biostatistics Working Group"
---

[*add diagrams/figures]

## Overview
Generalized estimating equations (GEEs) are a statistical method used to analyze longitudinal or clustered data, while accounting for within-subject correlation. The goal GEEs are to make inferences about population-averaged effects (controlling for within-subject correlations), rather than individual subject-level effects. GEEs can handle a wide range of outcome distributions, including binary, count, and continuous data (Fitzmaurice et al., 2011; Hardin & Hilbe, 2012). This method is an extension of the generalized linear model (Diggle et al., 2002), shows good statistical power and efficiency, is robust to non-normality and non-constant variance and can handle unbalanced or unequally spaced data.

### When to use Longitudinal Generalized Estimating Equations (GEEs)?
You should use xxxxxxx in the following scenario:
1.	You want to know xxxxx
2.	Your variable xxxxx
3.	You have xxxxx

### Getting Started with Longitudinal Generalized Estimating Equations (GEEs)

In this tutorial, we will introduce the concept of longitudinal generalized estimating equations (GEEs) and guide you through a simple example using a small dataset. By the end of this tutorial, you will be able to:

1. Understand the basic concepts of longitudinal GEEs.
2. Fit a GEE model using example data in R.
3. Interpret the results of the model.

### Basic Example

For this tutorial, we will use a simple dataset containing the following information for a group of individuals:

- Subject ID
- Time (in years)
- Binary outcome variable (e.g., presence or absence of a particular condition)

| Subject ID | Time | Outcome |
|------------|------|---------|
| 1          | 0    | 0       |
| 1          | 1    | 1       |
| 1          | 2    | 1       |
| 2          | 0    | 0       |
| 2          | 1    | 1       |

### Model Specification and Estimation

We will use the `geepack` package in R to fit a longitudinal GEE model. First, install and load the required package:

```{r}
if (!("geepack" %in% installed.packages())) install.packages("geepack")
library(geepack)
```
Next, create a data frame with the example data:

```{r}
data <- data.frame(
  subject_id = c(1, 1, 1, 2, 2),
  time = c(0, 1, 2, 0, 1),
  outcome = c(0, 1, 1, 0, 1)
)
```

Fit the longitudinal GEE model with a binary outcome using a logit link function and an exchangeable correlation structure:

```{r}
model <- geeglm(outcome ~ time, data = data, id = subject_id, family = binomial(link = "logit"), corstr = "exchangeable")
```

### Interpreting the Results
To interpret the results of the longitudinal GEE model, we can use the summary() function in R to display the estimated population-averaged effects:

```{r}
summary(model)
```
The output will show the estimated effect of time on the binary outcome variable, expressed as an odds ratio. By examining the odds ratio and its associated p-value, we can determine if the change in the outcome variable over time is significant at the population level.