### 2023-04-20 14_Tutorials_StateTraitModels.qmd

---
title: "State-Trait Models"
author: "Biostatistics Working Group"
---

## Overview
State-Trait Models are a type of statistical model used to separate the variance in a given variable into two components: a stable, trait-like component and a time-specific, state-like component. In this tutorial, we will walk you through a simple example using a simulated dataset and demonstrate how to fit a State-Trait Model using the `lavaan` package in R.

[+add diagrams/figures]

### When to use State-Trait Models?
You should use xxxxxxx in the following scenario:

1.	You want to know xxxxx
2.	Your variable xxxxx
3.	You have xxxxx

### Getting Started with State-Trait Models

#### Prerequisites
To follow along with this tutorial, you will need to have R and RStudio installed on your computer. Additionally, you will need to install the `lavaan` package, which provides functions for fitting various structural equation models, including State-Trait Models.

## Basic Example 
```{r setup, echo=TRUE, warning=FALSE, message=FALSE}
# Install the lavaan package if not already installed
if (!("lavaan" %in% installed.packages())) {
  install.packages("lavaan")
}
```

# Load the lavaan package
```{r}
library(lavaan)
```

## Simulating the Data
For this tutorial, we will use a simulated dataset with 300 individuals, each measured at four time points. The dataset will contain a variable with both state-like and trait-like components.

```{r}
set.seed(123)
n <- 300
timepoints <- 4

# Simulate the trait-like component
trait <- rnorm(n, mean = 100, sd = 10)

# Simulate the state-like component
state <- matrix(rnorm(n * timepoints, mean = 0, sd = 5), nrow = n, ncol = timepoints)

# Combine trait and state components
data <- data.frame(id = rep(1:n, each = timepoints),
                   time = rep(1:timepoints, times = n),
                   value = c(t(trait + state)))
```

### Model Specification and Estimation
Now that we have our simulated data, we can fit a State-Trait Model using the lavaan package. The model will estimate the trait-like and state-like components for each individual.

```{r}
# Define the model
model <- '
  # Trait component
  trait =~ 1 * value_t1 + 1 * value_t2 + 1 * value_t3 + 1 * value_t4

  # State component
  state =~ value_t1 + value_t2 + value_t3 + value_t4

  # Residual variances
  value_t1 ~~ value_t1
  value_t2 ~~ value_t2
  value_t3 ~~ value_t3
  value_t4 ~~ value_t4
'
```

```{r}
# Reshape the data to wide format
data_wide <- reshape(data, idvar = "id", timevar = "time", direction = "wide")
```

```{r, eval=FALSE}
# Fit the State-Trait Model
fit <- sem(model, data = data_wide)
```

### Interpreting the Results
Now that we have estimated our State-Trait Model, we can interpret the results. The key parameters of interest in this model are the trait-like and state-like components, which are the variances attributed to each component.

```{r, eval=FALSE}
# Obtain the results
summary(fit)
```