### 2023-04-20 15_Tutorials_RandomInterceptCrosslaggedPanelModels.qmd

---
title: "Random-Intercept Crosslagged Panel Models"
author: "Biostatistics Working Group"
---

## Overview
Random-Intercept Cross-Lagged Panel Models (RI-CLPM) are a type of structural equation model used to analyze longitudinal data with repeated measures of multiple variables. In this tutorial, we will walk you through a simple example using a simulated dataset and demonstrate how to fit a RI-CLPM using the `lavaan` package in R.

[+add diagrams/figures]

### When to use Random-Intercept Cross-Lagged Panel Models?
You should use xxxxxxx in the following scenario:

1.	You want to know xxxxx
2.	Your variable xxxxx
3.	You have xxxxx

### Getting Started with Random-Intercept Cross-Lagged Panel Models

#### Prerequisites
To follow along with this tutorial, you will need to have R and RStudio installed on your computer. Additionally, you will need to install the `lavaan` package, which provides functions for fitting various structural equation models, including RI-CLPMs.

```{r setup, echo=TRUE, warning=FALSE, message=FALSE}
# Install the lavaan package if not already installed
if (!("lavaan" %in% installed.packages())) {
  install.packages("lavaan")
}
# Load the lavaan package
library(lavaan)
```

## Basic Example 
Simulating the Data
For this tutorial, we will use a simulated dataset with 500 individuals, each measured at five time points. The dataset will contain two variables, X and Y, with cross-lagged effects.

```{r}
set.seed(123)
n <- 500
timepoints <- 5

# Simulate the data
data <- data.frame(id = rep(1:n, each = timepoints),
                   time = rep(1:timepoints, times = n),
                   X = rnorm(n * timepoints, mean = 0, sd = 1),
                   Y = rnorm(n * timepoints, mean = 0, sd = 1))

# Introduce cross-lagged effects
for (i in 2:timepoints) {
  data$X[data$time == i] <- 0.3 * data$Y[data$time == (i - 1)] + data$X[data$time == i]
  data$Y[data$time == i] <- 0.2 * data$X[data$time == (i - 1)] + data$Y[data$time == i]
}
```

### Model Specification and Estimation
Now that we have our simulated data, we can fit a Random-Intercept Cross-Lagged Panel Model using the lavaan package.

```{r, eval=FALSE}
# Define the model
model <- '
  # Random intercepts
  i_X =~ 1 * X_t1 + 1 * X_t2 + 1 * X_t3 + 1 * X_t4 + 1 * X_t5
  i_Y =~ 1 * Y_t1 + 1 * Y_t2 + 1 * Y_t3 + 1 * Y_t4 + 1 * Y_t5

  # Cross-lagged effects
  Y_t2 ~ beta_YX * X_t1
  Y_t3 ~ beta_YX * X_t2
  Y_t4 ~ beta_YX * X_t3
  Y_t5 ~ beta_YX * X_t4

  X_t2 ~ beta_XY * Y_t1
  X_t3 ~ beta_XY *
```

### Interpreting the Results
xxxxxx