### 2023-04-20 16_Tutorials_LatentCurveModelsStructuredResiduals.qmd

---
title: "Latent Curve Models with Structured Residuals"
author: "Biostatistics Working Group"
---

## Overview
In this tutorial, we will explore Latent Curve Models with Structured Residuals (LCM-SR). LCM-SR allows for the separation of between-person and within-person components of individual change over time. By using structured residuals, we can account for time-varying covariates and interactions between variables at both the within-person and between-person levels.

[+add diagrams/figures]

### When to use Latent Curve Models with Structured Residuals?
You should use xxxxxxx in the following scenario:

1.	You want to know xxxxx
2.	Your variable xxxxx
3.	You have xxxxx

### Getting Started with Latent Curve Models with Structured Residuals

#### Prerequisites
You need to have the `lavaan` and `semTools` packages installed in R to perform LCM-SR. Install them using the following commands:

```{r}
if (!("lavaan" %in% installed.packages())) install.packages("lavaan")
if (!("semTools" %in% installed.packages())) install.packages("semTools")
```

## Basic Example 
For this tutorial, we will use simulated data with 500 individuals and 5 timepoints. The data will include a time-varying covariate, covariate, and an outcome variable, outcome.

```{r}
set.seed(123)
library(lavaan)
library(semTools)

# Simulate data
n <- 500
timepoints <- 5

id <- rep(1:n, each = timepoints)
time <- rep(1:timepoints, n)

covariate <- rnorm(n * timepoints, mean = 0, sd = 1)
outcome <- rnorm(n * timepoints, mean = 0, sd = 1)

data <- data.frame(id, time, covariate, outcome)
```

### Model Specification and Estimation
To specify an LCM-SR, we will use the lavaan syntax to define the measurement model and the structural model. We will model the linear growth of the outcome variable while accounting for the effect of the time-varying covariate.

```{r}
model <- '
  # Latent variables
  intercept =~ 1 * outcome_t1 + 1 * outcome_t2 + 1 * outcome_t3 + 1 * outcome_t4 + 1 * outcome_t5
  slope =~ 0 * outcome_t1 + 1 * outcome_t2 + 2 * outcome_t3 + 3 * outcome_t4 + 4 * outcome_t5

  # Time-varying covariate effect on outcome
  outcome_t1 ~ c1 * covariate_t1
  outcome_t2 ~ c2 * covariate_t2
  outcome_t3 ~ c3 * covariate_t3
  outcome_t4 ~ c4 * covariate_t4
  outcome_t5 ~ c5 * covariate_t5

  # Residual variances and covariances
  outcome_t1 ~~ r1 * outcome_t1
  outcome_t2 ~~ r2 * outcome_t2
  outcome_t3 ~~ r3 * outcome_t3
  outcome_t4 ~~ r4 * outcome_t4
  outcome_t5 ~~ r5 * outcome_t5

  # Latent variable variances
  intercept ~~ i_var * intercept
  slope ~~ s_var * slope

  # Latent variable covariances
  intercept ~~ i_s_cov * slope
'
```

Model Estimation
Now we will estimate the model using the lavaan function sem().

```{r, eval=FALSE}
# Reshape data to wide format
wide_data <- spread(data, key = time, value = outcome, sep
```
### Interpreting the Results
xxxxx