### 2023-04-21 test.tex
cli commit 10

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

% load packages
\usepackage{geometry}
\usepackage{xcolor}
\usepackage{eso-pic}
\usepackage{fancyhdr}
\usepackage{sectsty}
\usepackage{fontspec}
\usepackage{titlesec}

%% Set page size with a wider right margin
\geometry{a4paper, total={170mm,257mm}, left=20mm, top=20mm, bottom=20mm, right=50mm}

%% Let's define some colours
\definecolor{light}{HTML}{E6E6FA}
\definecolor{highlight}{HTML}{800080}
\definecolor{dark}{HTML}{330033}

%% Let's add the border on the right hand side 
\AddToShipoutPicture{% 
    \AtPageLowerLeft{% 
        \put(\LenToUnit{\dimexpr\paperwidth-3cm},0){% 
            \color{light}\rule{3cm}{\LenToUnit\paperheight}%
          }%
     }%
     % logo
    \AtPageLowerLeft{% start the bar at the bottom right of the page
        \put(\LenToUnit{\dimexpr\paperwidth-2.25cm},27.2cm){% move it to the top right
            \color{light}\includegraphics[width=1.5cm]{_extensions/nrennie/PrettyPDF/logo.png}
          }%
     }%
}

%% Style the page number
\fancypagestyle{mystyle}{
  \fancyhf{}
  \renewcommand\headrulewidth{0pt}
  \fancyfoot[R]{\thepage}
  \fancyfootoffset{3.5cm}
}
\setlength{\footskip}{20pt}

%% style the chapter/section fonts
\chapterfont{\color{dark}\fontsize{20}{16.8}\selectfont}
\sectionfont{\color{dark}\fontsize{20}{16.8}\selectfont}
\subsectionfont{\color{dark}\fontsize{14}{16.8}\selectfont}
\titleformat{\subsection}
  {\sffamily\Large\bfseries}{\thesection}{1em}{}[{\titlerule[0.8pt]}]

% left align title
\makeatletter
\renewcommand{\maketitle}{\bgroup\setlength{\parindent}{0pt}
\begin{flushleft}
  {\sffamily\huge\textbf{\MakeUppercase{\@title}}} \vspace{0.3cm} \newline
  {\Large {\@subtitle}} \newline
  \@author
\end{flushleft}\egroup
}
\makeatother

%% Use some custom fonts
\setsansfont{Ubuntu}[
    Path=_extensions/nrennie/PrettyPDF/Ubuntu/,
    Scale=0.9,
    Extension = .ttf,
    UprightFont=*-Regular,
    BoldFont=*-Bold,
    ItalicFont=*-Italic,
    ]

\setmainfont{Ubuntu}[
    Path=_extensions/nrennie/PrettyPDF/Ubuntu/,
    Scale=0.9,
    Extension = .ttf,
    UprightFont=*-Regular,
    BoldFont=*-Bold,
    ItalicFont=*-Italic,
    ]
\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\@ifundefined{codebgcolor}{\definecolor{codebgcolor}{named}{light}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Longitudinal Analysis Manuscript: Working Draft},
  colorlinks=true,
  linkcolor={highlight},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={highlight},
  pdfcreator={LaTeX via pandoc}}

\title{Longitudinal Analysis Manuscript: Working Draft}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Longitudinal Analysis using data from the ABCD Study}
\author{Samuel W. Hawes \and Andrew K. Littlefield \and Daniel A.
Lopez \and Kenneth J. Sher \and Additional Co-authors \and Wesley K.
Thompson}
\date{}

\begin{document}
\maketitle
\begin{abstract}
The Adolescent Brain Cognitive Development (ABCD) Study presents a
unique opportunity for researchers to investigate developmental
processes in a large, diverse cohort of children and adolescents. Given
the complex nature of the longitudinal data collected in the ABCD Study,
researcher are likely to encounter a myriad of methodological and
analytic considerations and concerns. This review provides a
comprehensive examination of key issues and techniques related to
longitudinal data analysis, specifically focusing on the ABCD Study. The
text discusses model assumptions, common violations (e.g., independent
and identically distributed residuals, \emph{heterogeneous}
\emph{variability}) and their implications for valid inference. The
importance of appropriately modeling covariance structures,
understandings trade-offs between model fit and parsimony, and
challenges related to sample size, attrition, missing data are
highlighted. Consideration is given to the importance of selecting
appropriate statistical models to account for correlations in repeated
measurements and the assumptions underlying these models. The review
also differentiates between linear and non-linear models in the context
of continuous and discrete data, emphasizing various distributional
assumptions and the necessity of choosing appropriate models and
statistical methods. By addressing these complexities, the review seeks
to equip researchers with the necessary knowledge and tools to make
informed decisions as they navigate effectively analyzing and
interpreting data available in the ABCD Study.
\end{abstract}
\pagestyle{mystyle}

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, frame hidden, sharp corners, breakable, colback={codebgcolor}, enhanced, boxrule=0pt]}{\end{tcolorbox}}\fi

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\label{sec:headings} The Adolescent Brain Cognitive Development (ABCD)
Study® is the largest long-term investigation of neurodevelopment and
child health in the United States. Conceived and initiated by the
National Institutes of Health (NIH), this landmark prospective
longitudinal study aims to transform our understanding of the genetic
and environmental influences on brain development and their roles in
behavioral and health outcomes in adolescents (Volkow et al. 2018). At
its heart, the study is designed to chart the course of human
development across multiple, interacting domains from late childhood to
early adulthood and to identify factors that lead to both positive and
negative developmental outcomes. Central to achieving these goals is the
ABCD Study's® commitment to an open science framework designed to
facilitate access to and sharing of scientific knowledge by espousing
practices that increase openness, integrity, and reproducibility of
scientific research (e.g., public data releases). In this sense, the
ABCD Study® is a collaboration with the larger research community, with
the rich longitudinal nature of the ABCD Study dataset allowing
researchers to perform a variety of analyses of both methodological and
substantive interest. Together, this presents a unique opportunity to
significantly advance our understanding of how a multitude of
biopsychosocial processes emerge and unfold across critical periods of
development.

{[}section still be developed\ldots{]}

\hypertarget{the-abcd-study-data}{%
\subsection{The ABCD Study® Data}\label{the-abcd-study-data}}

Participants enrolled in the ABCD Study include a large cohort of youth
(n=11880) aged 9-10 years at baseline and their parents/guardians. The
study sample was recruited from household populations in defined
catchment areas for each of the 21 study sites across the United States
(information regarding funding agencies, recruitment sites,
investigators, and project organization can be obtained at the ABCD
Study website). The ABCD Study is collecting longitudinal data on a rich
variety of outcomes that will enable the construction of
realistically-complex etiological models by incorporating factors from
many domains simultaneously. Each new wave of data collection provides
the building blocks for conducting probing longitudinal analyses that
allow us to characterize normative development, identify variables that
presage deviations from prototypic development, and assess a range of
outcomes associated with variables of interest. This data includes a
neurocognitive battery (Luciana et al. 2018; Thompson et al. 2019),
mental and physical health assessments (Barch et al. 2018), measures of
culture and environment (Zucker et al. 2018), substance use {[}add
citation{]}, biospecimens (Uban et al. 2018), structural and functional
brain imaging (Casey et al. 2018; Hagler et al. 2019), geolocation-based
environmental exposure data, wearables, and mobile technology (Bagot et
al. 2018), and whole genome genotyping (Loughnan et al. 2020). Many of
these measures are collected at in-person annual visits, with brain
imaging collected at baseline and every other year going forward. A
limited number of assessments are collected in semi-annual telephone
interviews between in-person visits. Data are publicly released on an
annual basis through the NIMH Data Archive. By necessity, the study's
earliest data releases were cross-sectional (i.e., the baseline data),
however, the most recent public data release (NDA Release 4.0) contains
data collected across three annual assessments, including two imaging
assessments (baseline and year 2 follow-up visits).

\hypertarget{organization-of-current-manuscript}{%
\subsection{Organization of current
manuscript}\label{organization-of-current-manuscript}}

The rich longitudinal nature of the ABCD Study dataset will allow
researchers to perform analyses of both methodological and substantive
interest. This report describes methods for longitudinal analyses of
ABCD Study data that can address its fundamental scientific aims, as
well as challenges inherent in a large population-based long-term study
of adolescents. The manuscript is organized as follows:

{[}section still be developed\ldots{]}

\hypertarget{part-i-longitudinal-research-basic-concepts-and-considerations}{%
\section{Part I: Longitudinal Research: Basic Concepts and
Considerations}\label{part-i-longitudinal-research-basic-concepts-and-considerations}}

\label{sec:headings} There are several important concepts to consider
when conducting longitudinal analyses in a developmental context. These
include different ways of thinking about developmental course, whether
certain periods of development are relatively sensitive or insensitive
to various types of insults or stressors, whether some time periods or
situations inhibit the expression of individual differences due to
extreme environmental pressures, and whether the same behavior
manifested at different times represent the same phenomenon or different
ones. Further, in the case of developmentally focused longitudinal
research, each new measurement occasion not only provides a more
extended portrait of the child's life course (and not just characterize
growth during this period but also assesses the durability/chronicity of
prior effects/consequences) but also brings with it greater
methodological opportunities to exploit the statistical properties of
longitudinal data in the furtherance of critical scientific questions.
That is, we can ask more nuanced questions and make stronger inferences
as our number of time-ordered observations grows, assuming we have
assessed the ``right'' variables and the timings of our observations
comport with the temporal dynamics of the mechanisms of interest.
Appreciation of these and other issues can help to guide the analysis
and interpretation of data and aid translation to clinical and public
health applications.

\textbf{Vulnerable periods}. Development normatively progresses from
less mature to more mature levels of functioning. However, unique epochs
and experiences can alter the course of this idealized form of
development. Consider research that shows cannabis use during
adolescence is associated with later psychosis to a greater degree than
cannabis use initiated later in development {[}add citation{]}; or,
similarly, experimental research on rodents that shows rodent brains to
be especially sensitive to the neurotoxic effects of alcohol on brain
structure and learning early in development (corresponding to early
adolescence in humans){[}add citation{]}. These examples highlight the
importance of considering the role of vulnerable periods -- temporal
windows of rapid brain development or remodeling during which the
effects of environmental stimuli (e.g.~cannabis exposure) on the
developing brain may be particularly pronounced-- when trying to
establish an accurate understanding of the association between exposures
and outcomes.

\textbf{Developmental disturbances}. Whereas vulnerable periods heighten
neurobiological susceptibility to environmental influences, at other
times environmental pressures will tend to suppress stability and
disrupt the orderly stochastic process of normative development (e.g.,
xxx-xxx). This situation reflects a developmental disturbance in that
the normal course of development is ``disturbed'' for some time by some
time-limited process. In such cases, we might find that prediction of
behavior in the period of the disturbance is reduced and/or, similarly,
the behavior exhibited during the disturbance might have less predictive
power with respect to distal outcomes compared to the behavior exhibited
before and following the disrupted period. That is, once the
environmental stimuli are removed (or the individual is removed from the
environment), individual differences are again more freely expressed and
the autoregressive effects increase to levels similar to those before
entering the environment.

\textbf{Developmental snares and cascade effects}. Normative development
can also be upended by experiences (e.g., drug use) that, through
various mechanisms, disrupt the normal flow of development wherein each
stage establishes a platform for the next. For instance, substance use
could lead to association with deviant peers, precluding opportunities
for learning various adaptive skills and prosocial behaviors, in effect,
creating a ``snare'' that retards psychosocial development. Relatedly,
the consequences of these types of events can cascade (e.g., school
dropout, involvement in the criminal justice system) so that the effects
of the snare are amplified. Although conceptually distinct from
vulnerable periods, both of these types of developmental considerations
highlight the importance of viewing behavior in the context of
development and the importance of attempting to determine how various
developmental pathways unfold.

\textbf{Distinguishing developmental change from experience effects}.
One can often observe systematic changes over time in a variable of
interest and assume this change is attributable to development. To this
point, cognitive abilities (e.g, verbal ability, problem-solving)
normatively grow earlier in development and often decline in late life
(e.g., memory, speed of processing). However, the observed patterns of
growth and decline often differ between cross-sectional vs.~longitudinal
effects (Salthouse 2014) where subjects gain increasing experience with
the assessment with each successive measurement occasion. Such
experience effects on cognitive functioning have been demonstrated in
adolescent longitudinal samples similar to ABCD (Sullivan et al. 2017)
and highlight the need to consider these effects and address them
analytically. In the case of performance-based measures {[}e.g., matrix
reasoning related to neurocognitive functioning; see Salthouse
(2014){]}, this can be due to ``learning'' the task from previous test
administrations (e.g., someone taking the test a second time performs
better than they did the first time simply as a function of having taken
it before). Even in the case of non-performance-based measures (e.g.,
levels of depression), where one cannot easily make the argument that
one has acquired some task-specific skill through learning, it has been
observed that respondents tend to endorse lower levels on subsequent
assessments (e.g., Beck et al. 1961; see French and Sutton 2010) and
this phenomenon has been well documented in research on structured
diagnostic interviews (Robins 1985). While it is typically assumed that
individuals are rescinding or telling us less information on follow-up
interviews, there is reason to suspect that in some cases the initial
assessment may be artefactually elevated (see Shrout et al. 2018). Some
designs (specifically, accelerated longitudinal designs) are especially
well suited for discovering these effects and modeling them. While ABCD
was not designed as an accelerated longitudinal design, the variability
in age at the time of baseline recruitment (9 years, 0 months to 10
years, 11 months) allows some measures, collected every year, to be
conceptualized as an accelerated longitudinal design. Moreover, it is
possible that in later waves, patterns of longitudinal missing data will
allow some analyses to assess the confounded effects of age and the
number of prior assessments. However, ABCD is fundamentally a
single-cohort, longitudinal design, where a number of prior assessments
and age are highly confounded, and for, perhaps, most analyses, the
possible influence of experience effects needs to be kept in mind.

\hypertarget{part-ii-longitudinal-data-interpretation-issues-pitfalls-assumption}{%
\section{Part II Longitudinal Data: Interpretation / Issues / Pitfalls
\&
Assumption}\label{part-ii-longitudinal-data-interpretation-issues-pitfalls-assumption}}

\label{sec:headings} \textbf{Defining Features of Longitudinal Data
Analysis.} The hallmark characteristic of longitudinal data analysis is
its application to repeated assessments of the same assessment targets
(e.g., individuals, families) across time. While the primary reason for
collecting longitudinal data is in pursuit of addressing scientific
questions, from a methodological perspective, having multiple
observations over time allows researchers to identify potentially
problematic observations when highly improbable longitudinal patterns
are observed. That is, we can ask more nuanced questions and make
stronger inferences as our number of time-ordered observations grows
assuming we have assessed the ``right'' variables and the timings of our
observations comport with the temporal dynamics of the mechanisms of
interest .

\hypertarget{modeling-data-across-two-time-points-versus-three-or-more-time-points.}{%
\subsection{Modeling Data Across Two Time Points versus Three or More
Time
Points.}\label{modeling-data-across-two-time-points-versus-three-or-more-time-points.}}

Although the clear leap to the realm of longitudinal data involves going
from one assessment to two or more assessments, there are also notable
distinctions in designs based on two-assessment points versus three or
more measurement occasions. Just as cross-sectional data can be
informative in some situations, two waves of data can be beneficial in
contexts such as when experimental manipulation is involved (e.g.,
pre/post tests), or if the central goal is prediction (e.g., trying to
predict scores on Variable A at time T as a function of prior scores on
Variable A and Variable B at time T-1). At the same time, data based on
two assessments are inherently limited on multiple fronts. As (Rogosa,
Brandt, and Zimowski 1982) noted approximately forty years ago, ``Two
waves of data are better than one, but maybe not much better''. These
sentiments are reflected in more contemporary recommendations regarding
best-practice guidelines for prospective data, which increasingly
emphasize the benefits of additional measurement occasions for model
identification and accurate parameter estimation. It is also consistent
with research recommending that developmental studies include three or
more assessment points, given it is impossible for data based on
two-time points to determine the shape of development (since linear,
straight line change is the only possible form, given two assessments;
see (Duncan and Duncan 2009)). Research designs that include three or
more time points allow for increasingly nuanced analyses that more
adequately tease apart sources of variation and covariation among the
repeated assessments (King et al. 2018)-- a key aspect of inferential
research. To illustrate, developmental theories are typically interested
in understanding patterns of within-individual change over time
(discussed in further detail, below); however, two data points provide
meager information on change at the person level. This point is further
underscored in a recent review of statistical models commonly touted as
distinguishing within-individual vs between-individual sources of
variance in which the study authors concluded ``\ldots{} researchers are
limited when attempting to differentiate these sources of variation in
psychological phenomenon when using two waves of data'' and perhaps more
concerning, ``\ldots the models discussed here do not offer a feasible
way to overcome these inherent limitations'' Littlefield et al. (2021).
It is important to note, however, that despite the current focus on
two-wave designs versus three or more assessment waves, garnering three
assessment points is not a panacea for longitudinal modeling. Indeed,
several contemporary longitudinal models designed to isolate
within-individual variability {[}e.g., the Latent Curve Model with
Structured Residuals; Curran et al. (2014){]} require at least four
assessments to parameterize fully and, more generally, increasingly
accurate parameter estimates are obtained as more assessment occasions
are used (Duncan and Duncan 2009).

\hypertarget{types-of-stability-and-change}{%
\subsection{Types of stability and
change}\label{types-of-stability-and-change}}

If one were to try to sum up what development in a living organism is
exactly, one could plausibly argue it's the characterization of
stability and change as the organism traverses the life course. There
are a few different ways to think of stability (and change). Consider we
measure the height of all youth in a 6th-grade class, once in the fall
at the beginning of the school year and once again in the spring at the
end of the school year. A common first step may be to compare the
class's average height values obtained at these two different
measurement occasions. This comparison of the average scores for the
same group of individuals at multiple time points is referred to as
``mean-level'' stability as it provides information about continuity and
change in the group level of an outcome of interest (e.g., height) over
time. Another type of stability involves calculating the correlation
between the values obtained at different time points (e.g., `height in
the fall' with `height in the spring'). This type of ``rank-order''
stability evaluates between-individual change by focusing on the degree
to which individuals retain their relative placement in a group across
time. Consider, someone who is the shortest person in their class in 6th
grade may grow considerably over the school year (i.e., exhibit mean
level change), but remain the shortest person among their classmates.
That is, the individual is manifesting a type of rank-order stability.
Both types of stability and change are important. Mean-level change in
certain traits might help to explain why, in general, individuals are
particularly vulnerable to social influences at some ages more than
others; rank order change might help to quantify the extent to which
certain characteristics of the individual are more trait-like. For
example, in some areas of development, there is considerable mean-level
change that occurs over time (e.g., changes in Big 5 personality
traits), but relatively high rank-order stability. Despite the useful
information afforded by examining mean-level and rank-order change,
these approaches are limited in that they provide little information
about patterns of ``within-individual'' change and, in turn, can result
in fundamental misinterpretations about substantial or meaningful
changes in an outcome of interest.

There is growing recognition that statistical models commonly applied to
longitudinal data often fail to comport with the developmental theory
they are being used to assess (e.g., Curran, Lee, Howard, Lane, \&
MacCallum, 2012; Hoffman, 2015; Littlefield et al., 2021. Specifically,
developmental studies typically involve the use of prospective data to
inform theories that are concerned with clear within-person (i.e.,
intraindividual) processes (e.g., how phenotypes change or remain stable
within individuals over time) (e.g., see Curran and Bauer 2011). Despite
this, methods generally unsuited for disaggregating between- and
within-person effects (e.g., cross-lagged panel models {[}CLPM{]})
remain common within various extant literatures. As a result, experts
increasingly caution about the need to xxxxxxxx {[}add citation{]}.
Fortunately, there exists a range of models that have been proposed to
tease apart between- and within-person sources of variance across time
(see Littlefield et al. 2021; Orth et al. 2021). Most of these
contemporary alternatives incorporate time-specific latent variables to
capture between-person sources of variance and model within-person
deviations around an individual's mean (or trait) level across time
(e.g., RI-CLPM, Hamaker, Kuiper, and Grasman 2015; LCM-SR, Curran et al.
2014). It is important to note however that these models require
multiple assessments waves (e.g., four or more to fully specify the
LCM-SR), additional expertise to overcome issues with model convergence,
and appreciation of modeling assumptions when attempting to adjudicate
among potential models in each research context (see Littlefield et al.
2021, for further discussion).

\hypertarget{model-assumptions}{%
\subsection{Model Assumptions}\label{model-assumptions}}

Many statistical models assume certain characteristics about the data to
which they are being applied. As an example, common assumptions of
parametric statistical models include normality, linearity, and equality
of variances. These assumptions must be carefully considered before
conducting analysis so that valid inferences can be made from the data;
that is, violation of a model's assumptions can substantively alter the
interpretation of results. Similarly, statistical models employed in the
analyses of longitudinal data often entail a range of assumptions that
must be closely inspected. One central issue for repeated measurements
on an individual is how to account for the correlated nature of the
data; another common feature of longitudinal data is heterogeneous
variability; that is, the variance of the response changes over the
duration of the study. Traditional techniques, such as a standard
regression or ANOVA model, assume residuals are independent and thus are
inappropriate for designs that assess (for example) the same individuals
across time. That is, given the residuals are no longer independent, the
standard errors from the models are biased and can produce misleading
inferential results. Although there are formal tests of independence for
time series data (e.g., the Durbin-Watson statistic (Durbin and Watson
1950)), more commonly independence is assumed to be violated in study
designs with repeated assessments. Therefore, an initial question to be
addressed by a researcher analyzing prospective data is how to best
model the covariance structure of said data.

\hypertarget{covariance-structures}{%
\subsection{Covariance Structures}\label{covariance-structures}}

Statistical models for longitudinal data include two main components to
account for assumptions that are commonly violated when working with
repeated measures data: a model for the covariance among repeated
measures (both the correlations among pairs of repeated measures on an
individual and the variability of the responses on different occasions),
coupled with a model for the mean response and its dependence on
covariates (eg, treatment group in the context of clinical trials). This
allows for the specification of a range of so-called covariance
structures, each with its own set of tradeoffs between model fit and
parsimony (e.g., see Kincaid 2005).

\hypertarget{accounting-for-correlated-data}{%
\subsection{Accounting for Correlated
Data}\label{accounting-for-correlated-data}}

As an example, one alternative structure that attempts to handle the
reality that correlations between repeated assessments tend to diminish
across time is the autoregressive design. As the name implies, the
structure assumes a subsequent measurement occasion (e.g., assessment at
Wave 2) is regressed onto (that is, is predicted by) a prior measurement
occasion (e.g., assessment at Wave 1). The most common type of
autoregressive design is the AR(1), where assessments at time T + 1 are
regressed on assessments at Time T. Identical to compound symmetry, this
model assumes the variances are homogenous across time. Diverting from
compound symmetry, this model assumes the correlations between repeated
assessments decline exponentially across time rather than remaining
constant. For example, per the AR(1) structure, if the correlation
between Time 1 and Time 2 data is thought to be .5, then the correlation
between Time 1 and Time 3 data would be assumed to be .5\emph{.5 = .25,
and the correlation between Time 1 and Time 4 data would be assumed to
be .5}.5*.5 = .125. As with compound symmetry, the basic AR(1) model is
parsimonious in that it only requires two parameters (the variance of
the assessments and the autoregressive coefficient). Notably, the
assumption of constant autoregressive relations between assessments is
often relaxed in commonly employed designs that use autoregressive
modeling (e.g., cross-lagged panel models {[}CLPM{]}). These designs
still typically assume an AR(1) process (e.g., it is sufficient to
regress the Time 3 assessment onto the Time 2 assessment and is not
necessary to also regress the Time 3 assessment onto the Time 1
assessment, which would result in an AR(2) process). However, the
magnitude of these relations is often allowed to differ across different
AR(1) pairs of assessment (e.g., the relation between Time 1 and Time 2
can be different from the relation between Time 2 and Time 3). These
more commonly employed models also often relax the assumption of equal
variances of the repeated assessments. Although the AR(1) structure may
involve a more realistic set of assumptions compared to compound
symmetry, in that the AR(1) model allows for diminishing correlations
across time, the basic AR(1) model, as well as autoregressive models
more generally, can also suffer from several limitations in contexts
that are common in prospective designs. In particular, recent work
demonstrates that if a construct being assessed prospectively across
time is trait-like in nature, then autoregressive relations fail to
adequately account for this trait-like structure, with the downstream
consequence that estimates derived from models based on AR structures
(such as the CLPM) can be misleading and fail to adequately demarcate
between- vs.~within-person sources of variance (Hamaker, Kuiper, and
Grasman 2015).

\hypertarget{linear-vs-non-linear-models}{%
\subsection{Linear vs non-linear
models}\label{linear-vs-non-linear-models}}

Identification of optimal statistical models and appropriate
mathematical functions requires an understanding of the type of data
being used. Repeated assessments can be based on either continuous or
discrete measures. Examples of discrete measures include repeated
assessments of binary variables (e.g., past 12-month alcohol use
disorder status measured across ten years), ordinal variables (e.g., a
single item measuring the level of agreement to a statement on a
three-point scale including the categories of ``disagree'', ``neutral'',
and ``agree'' in an ecological momentary assessment study that involves
multiple daily assessments), and count variables (e.g., number of
cigarettes smoked per day across a daily diary study). In many ways, the
distributional assumptions of indicators used in longitudinal designs
mirror the decision points and considerations when delineating across
different types of discrete outcome variables, a topic that spans entire
textbooks (e.g., see Lenz 2016). For example, the Mplus manual (Muthén
2017) includes examples of a) censored and censored-inflated models, b)
linear growth models for binary or ordinal variables, c) linear growth
models for a count outcome assuming a Poisson model, d) linear growth
models for a count outcome assuming a zero-inflated Poisson model and e)
discrete- and continuous-time survival analysis for a binary outcome.
Beyond these highlighted examples, other distributions (e.g., negative
binomial) can be assumed for the indicators when modeling longitudinal
data. These models can account for issues that can occur when working
with discrete outcomes, including overdispersion (when the variance is
higher than would be expected based on a given distribution) and
zero-inflation {[}when more zeros occur than is expected based on a
given distribution; see Lenz (2016){]}. Models involving zero-inflation
parameters are referred to as two-part models, given one part of the
model predicts the zero-inflation whereas the other part of the model
predicts outcomes consistent with a given distribution {[}e.g., Poisson
distribution; see Farewell et al. (2017), for a review of two-part
models for longitudinal data{]}. Although there exist several
alternative models for discrete indicators, some more recent models that
have been proposed for prospective data are only feasible in cases where
indicators are assumed to be continuous rather than discrete {[}e.g.,
LCM-SR; Curran et al. (2014){]}. Given the sheer breadth of issues
relevant to determining better models for discrete outcomes, it is not
uncommon for texts on longitudinal data analysis to only cover models
and approaches that assume continuous indicators (e.g., T. D. Little
2013). However, some textbooks on categorical data analysis provide more
detailed coverage of the myriad issues and modeling choices to consider
when working with discrete outcomes {[}e.g., Lenz (2016), Chapter 11 for
matched pair/two-assessment designs; Chapter 12 for marginal and
transitional models for repeated designs, such as generalized estimating
equations, and Chapter 13 for random effects models for discrete
outcomes{]}.

\hypertarget{missing-dataattrition}{%
\subsection{Missing Data/Attrition}\label{missing-dataattrition}}

As recently reviewed by Littlefield (in press), investigators of
prospective data are confronted with study attrition (i.e., participants
may not provide data at a given wave of assessment) and thus approaches
are needed to confront the issue of missing data. Three models of
missingness are typically considered in the literature (see R. J. Little
and Rubin 1989). These three models are data: a) missing completely at
random (MCAR), b) missing at random (MAR), and c) missing not at random
(MNAR). Data that are MCAR means missing data is a random sample of all
the types of participants (e.g., males) in a given dataset. MAR suggests
conditionally missing at random (see Graham 2009). That is, MAR implies
missingness is completely random (i.e., does not hinge on some
unmeasured variables) once missingness has been adjusted by all
available variables in a dataset (e.g., biological sex). Data that are
MNAR are missing as a function of unobserved variables. Graham (2009)
provides an excellent and easy-to-digest overview of further details
involving missing data considerations.

Multiple approaches have been posited to handle missing data. Before the
advent of more contemporary approaches, common methods included several
ad hoc procedures. These include eliminating the data of participants
with missing data (e.g., listwise or pairwise deletion) or using mean
imputation (i.e., replacing the missing value with the mean score of the
sample that did participate). However, these methods are not recommended
because they can contribute to biased parameter estimates and research
conclusions (see Graham 2009). More specifically, the last observation
carried forward (LOCF) is a common approach to imputing missing data.
LOCF replaces a participant's missing values after dropout with the last
available measurement (Molnar, Hutton, and Fergusson 2008). This
approach assumes stability (i.e., a given participant's score is not
anticipated to increase or decline after study dropout) and that the
data are MCA R. However, as described by Molnar, Hutton, and Fergusson
(2008), it is common for treatment groups to show higher attrition
compared to control groups in studies of dementia drugs. Given that
dementia worsens over time, using LOCF biases the results in favor of
the treatment group (see Molnar, Hutton, and Fergusson 2008, for more
details).

More modern approaches, such as using maximum likelihood or multiple
imputation to estimate missing data, are thought to avoid some of the
biases of older approaches (see Enders 2010; Graham 2009). Graham (2009)
noted several ``myths'' regarding missing data. For example, Graham
notes many assume the data must be minimally MAR to permit estimating
procedures (such as maximum likelihood or multiple imputation) compared
to other, more traditional approaches (e.g., using only complete case
data). Violations of MAR impact both traditional and more modern data
estimation procedures, though as noted by Graham, violations of MAR tend
to have a greater effect on older methods. Graham thus suggests that
estimating missing data is a better approach compared to the older
procedures in most circumstances, regardless of the model of missingness
{[}i.e., MCAR, MAR, MNAR; see Graham (2009){]}.

Attrition from a longitudinal panel study such as ABCD is inevitable and
represents a threat to the validity of longitudinal analyses and
cross-sectional analyses conducted at later time points, especially
since attrition can only be expected to grow over time. While, to date,
attrition in ABCD has been minimal (some cite here), it remains an
important focus for longitudinal analysis and its significance is likely
to only grow as the cohort ages. Ideally, one tries to minimize
attrition through good retention practices from the outset via
strategies designed to maintain engagement in the project (Cotter et al.
2005; Hill et al. 2016; Watson et al. 2018). However, even the
best-executed studies need to anticipate growing attrition over the
length of the study and implement analytic strategies designed to
provide the most valid inferences. Perhaps the most key concern when
dealing with data that is missing due to attrition is determining the
degree of bias in retained variables that is a consequence of attrition.
Assuming that the data are not missing completely at random, attention
to the nature of the missingness and employing techniques designed to
mitigate attrition-related biases need to be considered in all
longitudinal analyses. Several different approaches can be considered
and employed depending upon the nature of the intended analyses, the
degree of missingness, and data available to help estimate missing and
unobserved values.

\hypertarget{quantifying-effect-sizes-longitudinally}{%
\subsection{Quantifying effect sizes
longitudinally}\label{quantifying-effect-sizes-longitudinally}}

Given longitudinal data involve different sources of variance,
quantifying effect sizes longitudinally is a more difficult task
compared to deriving such estimates from cross-sectional data. Effect
size can be defined as, ``a population parameter (estimated in a sample)
encapsulating the practical or clinical importance of a phenomenon under
study.'' (Kraemer 2014). Common effect size metrics include r (i.e., the
standardized covariance, or correlation, between two variables) and
Cohen's d (Cohen 1988). Adjustments to common effect size calculations,
such as Cohen's d, are required even when only two time points are
considered (e.g., see Morris and DeShon 2002). Wang et al. (2019) note
there are multiple approaches to obtaining standardized within-person
effects, and that commonly suggested approaches (e.g., global
standardization) can be problematic (see Wang et al. 2019, for more
details). Thus, obtaining effect size metrics based on standardized
estimates that are relatively simple in cross-sectional data (such as r)
becomes more complex in the context of prospective data. Feingold (2009)
noted that equations for effects sizes used in studies involving growth
modeling analysis (e.g., latent growth curve modeling) were not
mathematically equivalent, and the effect sizes were not in the same
metric as effect sizes from traditional analysis (see Feingold 2009, for
more details). Given this issue, there have been various proposals for
adjusting effect size measures in repeated assessments. Feingold (2019)
reviews the approach for effect size metrics for analyses based on
growth modeling, including when considering linear and non-linear (i.e.,
quadratic) growth factors. Morris and DeShon (2002) review various
equations for effect size calculations relevant to when combining
estimates in meta-analysis with repeated measures and independent-groups
designs. Other approaches to quantifying effect sizes longitudinally may
be based on standardized estimates from models that more optimally
disentangle between- and within-person sources of variance (as reviewed
above). As an example, within a RI-CLPM framework, standardized
estimates between random intercepts (i.e., the correlation between two
random intercepts for two different constructs assessed repeatedly)
could be used to index the between-person relation, whereas standardized
estimates among the structured residuals could be used as informing the
effect sizes of within-person relations.

\hypertarget{part-iii-section-in-progress-tbd}{%
\section{Part III Section in Progress:
TBD}\label{part-iii-section-in-progress-tbd}}

\label{sec:headings} \textbf{Subsection TBD.}

\hypertarget{part-iv-discussion}{%
\section{Part IV Discussion}\label{part-iv-discussion}}

\label{sec:headings} \textbf{Subsection TBD.}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-bagot2018}{}}%
Bagot, K. S., S. A. Matthews, M. Mason, Lindsay M. Squeglia, J. Fowler,
K. Gray, M. Herting, A. May, Ian Colrain, and J. Godino. 2018.
{``Current, Future and Potential Use of Mobile and Wearable Technologies
and Social Media Data in the {ABCD} Study to Increase Understanding of
Contributors to Child Health.''} \emph{Developmental Cognitive
Neuroscience} 32: 121--29.

\leavevmode\vadjust pre{\hypertarget{ref-barch2018}{}}%
Barch, Deanna M., Matthew D. Albaugh, Shelli Avenevoli, Linda Chang,
Duncan B. Clark, Meyer D. Glantz, James J. Hudziak, Terry L. Jernigan,
Susan F. Tapert, and Debbie Yurgelun-Todd. 2018. {``Demographic,
Physical and Mental Health Assessments in the Adolescent Brain and
Cognitive Development Study: {Rationale} and Description.''}
\emph{Developmental Cognitive Neuroscience} 32: 55--66.

\leavevmode\vadjust pre{\hypertarget{ref-beck1961}{}}%
Beck, Aaron T., Calvin H. Ward, Mock Mendelson, Jeremiah Mock, and John
Erbaugh. 1961. {``An Inventory for Measuring Depression.''}
\emph{Archives of General Psychiatry} 4 (6): 561--71.

\leavevmode\vadjust pre{\hypertarget{ref-casey2018}{}}%
Casey, B. J., Tariq Cannonier, May I. Conley, Alexandra O. Cohen, Deanna
M. Barch, Mary M. Heitzeg, Mary E. Soules, et al. 2018. {``The
{Adolescent} {Brain} {Cognitive} {Development} ({ABCD}) Study: {Imaging}
Acquisition Across 21 Sites.''} \emph{Developmental Cognitive
Neuroscience}, The {Adolescent} {Brain} {Cognitive} {Development}
({ABCD}) {Consortium}: {Rationale}, {Aims}, and {Assessment} {Strategy},
32 (August): 43--54. \url{https://doi.org/10.1016/j.dcn.2018.03.001}.

\leavevmode\vadjust pre{\hypertarget{ref-cohen1988}{}}%
Cohen, Jacob. 1988. {``Statistical Power.''} \emph{Analysis for the
Behavioral Sciences}, 273--406.

\leavevmode\vadjust pre{\hypertarget{ref-cotter2005}{}}%
Cotter, Robert B., Jeffrey D. Burke, Magda Stouthamer-Loeber, and Rolf
Loeber. 2005. {``Contacting Participants for Follow-up: How Much Effort
Is Required to Retain Participants in Longitudinal Studies?''}
\emph{Evaluation and Program Planning} 28 (1): 15--21.

\leavevmode\vadjust pre{\hypertarget{ref-curran2011}{}}%
Curran, Patrick J., and Daniel J. Bauer. 2011. {``The Disaggregation of
Within-Person and Between-Person Effects in Longitudinal Models of
Change.''} \emph{Annual Review of Psychology} 62: 583--619.

\leavevmode\vadjust pre{\hypertarget{ref-curran2014a}{}}%
Curran, Patrick J., Andrea L. Howard, Sierra Bainter, Stephanie T. Lane,
and James S. McGinley. 2014. {``The {Separation} of {Between}-Person and
{Within}-Person {Components} of {Individual} {Change} {Over} {Time}: {A}
{Latent} {Curve} {Model} with {Structured} {Residuals}.''} \emph{J
Consult Clin Psychol} 82 (5): 879--94.
\url{https://doi.org/10.1037/a0035297}.

\leavevmode\vadjust pre{\hypertarget{ref-duncan2009}{}}%
Duncan, Terry E., and Susan C. Duncan. 2009. {``The {ABC}'s of {LGM}:
{An} {Introductory} {Guide} to {Latent} {Variable} {Growth} {Curve}
{Modeling}.''} \emph{Social and Personality Psychology Compass} 3 (6):
979--91. \url{https://doi.org/10.1111/j.1751-9004.2009.00224.x}.

\leavevmode\vadjust pre{\hypertarget{ref-durbin1950}{}}%
Durbin, James, and Geoffrey S. Watson. 1950. {``Testing for Serial
Correlation in Least Squares Regression: {I}.''} \emph{Biometrika} 37
(3/4): 409--28.

\leavevmode\vadjust pre{\hypertarget{ref-enders2010}{}}%
Enders, Craig K. 2010. \emph{Applied {Missing} {Data} {Analysis}}.
Guilford Press.

\leavevmode\vadjust pre{\hypertarget{ref-farewell2017}{}}%
Farewell, V. T., D. L. Long, B. D. M. Tom, S. Yiu, and L. Su. 2017.
{``Two-{Part} and {Related} {Regression} {Models} for {Longitudinal}
{Data}.''} \emph{Annual Review of Statistics and Its Application} 4 (1):
283--315.
\url{https://doi.org/10.1146/annurev-statistics-060116-054131}.

\leavevmode\vadjust pre{\hypertarget{ref-feingold2009}{}}%
Feingold, Alan. 2009. {``Effect Sizes for Growth-Modeling Analysis for
Controlled Clinical Trials in the Same Metric as for Classical
Analysis.''} \emph{Psychological Methods} 14 (1): 43.

\leavevmode\vadjust pre{\hypertarget{ref-feingold2019}{}}%
---------. 2019. {``Time-Varying Effect Sizes for Quadratic Growth
Models in Multilevel and Latent Growth Modeling.''} \emph{Structural
Equation Modeling: A Multidisciplinary Journal} 26 (3): 418--29.

\leavevmode\vadjust pre{\hypertarget{ref-french2010}{}}%
French, David P., and Stephen Sutton. 2010. {``Reactivity of Measurement
in Health Psychology: How Much of a Problem Is It? {What} Can Be Done
about It?''} \emph{British Journal of Health Psychology} 15 (3):
453--68.

\leavevmode\vadjust pre{\hypertarget{ref-graham2009}{}}%
Graham, John W. 2009. {``Missing {Data} {Analysis}: {Making} {It} {Work}
in the {Real} {World}.''} \emph{Annual Review of Psychology} 60 (1):
549--76. \url{https://doi.org/10.1146/annurev.psych.58.110405.085530}.

\leavevmode\vadjust pre{\hypertarget{ref-hagler2019}{}}%
Hagler, Donald J., SeanN. Hatton, M. Daniela Cornejo, Carolina Makowski,
Damien A. Fair, Anthony Steven Dick, Matthew T. Sutherland, et al. 2019.
{``Image Processing and Analysis Methods for the {Adolescent} {Brain}
{Cognitive} {Development} {Study}.''} \emph{NeuroImage} 202 (November):
116091. \url{https://doi.org/10.1016/j.neuroimage.2019.116091}.

\leavevmode\vadjust pre{\hypertarget{ref-hamaker2015}{}}%
Hamaker, Ellen L., Rebecca M. Kuiper, and Raoul P. P. P. Grasman. 2015.
{``A Critique of the Cross-Lagged Panel Model.''} \emph{Psychological
Methods} 20 (1): 102--16. \url{https://doi.org/10.1037/a0038889}.

\leavevmode\vadjust pre{\hypertarget{ref-hill2016}{}}%
Hill, Karl G., Danielle Woodward, Tiffany Woelfel, J. David Hawkins, and
Sara Green. 2016. {``Planning for Long-Term Follow-up: {Strategies}
Learned from Longitudinal Studies.''} \emph{Prevention Science} 17 (7):
806--18.

\leavevmode\vadjust pre{\hypertarget{ref-kincaid2005}{}}%
Kincaid, C. 2005. {``Guidelines for Selecting the Covariance Structure
in Mixed Model Analysis, Paper 198-30 in {Proceedings} of the
{Thirtieth} {Annual} {SAS} {Users} {Group} {Conference}.''} \emph{Inc.,
Cary, North Carolina}.

\leavevmode\vadjust pre{\hypertarget{ref-king2018}{}}%
King, Kevin M., Andrew K. Littlefield, Connor J. McCabe, Kathryn L.
Mills, John Flournoy, and Laurie Chassin. 2018. {``Longitudinal Modeling
in Developmental Neuroimaging Research: {Common} Challenges, and
Solutions from Developmental Psychology.''} \emph{Developmental
Cognitive Neuroscience}, Methodological {Challenges} in {Developmental}
{Neuroimaging}: {Contemporary} {Approaches} and {Solutions}, 33
(October): 54--72. \url{https://doi.org/10.1016/j.dcn.2017.11.009}.

\leavevmode\vadjust pre{\hypertarget{ref-kraemer2014}{}}%
Kraemer, Helena Chmura. 2014. {``Effect Size.''} \emph{The Encyclopedia
of Clinical Psychology}, 1--3.

\leavevmode\vadjust pre{\hypertarget{ref-lenz2016}{}}%
Lenz, Sylvia Tamara. 2016. {``Alan {Agresti} (2013): {Categorical} Data
Analysis.''} \emph{Statistical Papers} 57 (3): 849.

\leavevmode\vadjust pre{\hypertarget{ref-little1989}{}}%
Little, Roderick J., and Donald B. Rubin. 1989. {``The {Analysis} of
{Social} {Science} {Data} with {Missing} {Values}.''} \emph{Sociological
Methods \& Research} 18 (2-3): 292--326.
\url{https://doi.org/10.1177/0049124189018002004}.

\leavevmode\vadjust pre{\hypertarget{ref-little2013}{}}%
Little, Todd D. 2013. \emph{The {Oxford} {Handbook} of {Quantitative}
{Methods}, {Vol}. 2: {Statistical} {Analysis}}. Oxford University Press.

\leavevmode\vadjust pre{\hypertarget{ref-littlefield2021}{}}%
Littlefield, Andrew K., Kevin M. King, Samuel F. Acuff, Katherine T.
Foster, James G. Murphy, and Katie Witkiewitz. 2021. {``Limitations of
Cross-Lagged Panel Models in Addiction Research and Alternative Models:
{An} Empirical Example Using Project {MATCH}.''} \emph{Psychology of
Addictive Behaviors}. \url{https://doi.org/10.1037/adb0000750}.

\leavevmode\vadjust pre{\hypertarget{ref-loughnan2020}{}}%
Loughnan, Robert J., Clare E. Palmer, Wesley K. Thompson, Anders M.
Dale, Terry L. Jernigan, and Chun Chieh Fan. 2020. {``Polygenic Score of
Intelligence Is More Predictive of Crystallized Than Fluid Performance
Among Children.''} \emph{bioRxiv}, 637512.

\leavevmode\vadjust pre{\hypertarget{ref-luciana2018a}{}}%
Luciana, M., J. M. Bjork, B. J. Nagel, D. M. Barch, R. Gonzalez, S. J.
Nixon, and M. T. Banich. 2018. {``Adolescent Neurocognitive Development
and Impacts of Substance Use: {Overview} of the Adolescent Brain
Cognitive Development ({ABCD}) Baseline Neurocognition Battery.''}
\emph{Developmental Cognitive Neuroscience} 32: 67--79.

\leavevmode\vadjust pre{\hypertarget{ref-molnar2008}{}}%
Molnar, Frank J., Brian Hutton, and Dean Fergusson. 2008. {``Does
Analysis Using {`Last Observation Carried Forward'} Introduce Bias in
Dementia Research?''} \emph{Cmaj} 179 (8): 751--53.

\leavevmode\vadjust pre{\hypertarget{ref-morris2002}{}}%
Morris, Scott B., and Richard P. DeShon. 2002. {``Combining Effect Size
Estimates in Meta-Analysis with Repeated Measures and Independent-Groups
Designs.''} \emph{Psychological Methods} 7 (1): 105.

\leavevmode\vadjust pre{\hypertarget{ref-muthen2017}{}}%
Muthén, L. K. 2017. {``Mplus User's Guide. {Los} {Angeles}: {Muthén} \&
{Muthén}; 1998.''}

\leavevmode\vadjust pre{\hypertarget{ref-orth2021}{}}%
Orth, Ulrich, D. Angus Clark, M. Brent Donnellan, and Richard W. Robins.
2021. {``Testing Prospective Effects in Longitudinal Research:
{Comparing} Seven Competing Cross-Lagged Models.''} \emph{Journal of
Personality and Social Psychology} 120 (4): 1013.

\leavevmode\vadjust pre{\hypertarget{ref-robins1985}{}}%
Robins, Lee. 1985. {``Epidemiology: {Reflections} on {Testing} the
{Validity} of {Psychiatric} {Interviews} {\textbar} {JAMA} {Psychiatry}
{\textbar} {JAMA} {Network}.''}
\url{https://jamanetwork.com/journals/jamapsychiatry/article-abstract/493658}.

\leavevmode\vadjust pre{\hypertarget{ref-rogosa1982}{}}%
Rogosa, David, David Brandt, and Michele Zimowski. 1982. {``A Growth
Curve Approach to the Measurement of Change.''} \emph{Psychological
Bulletin} 92 (3): 726.

\leavevmode\vadjust pre{\hypertarget{ref-salthouse2014}{}}%
Salthouse, Timothy A. 2014. {``Why Are There Different Age Relations in
Cross-Sectional and Longitudinal Comparisons of Cognitive
Functioning?''} \emph{Current Directions in Psychological Science} 23
(4): 252--56.

\leavevmode\vadjust pre{\hypertarget{ref-shrout2018a}{}}%
Shrout, Patrick E., Gertraud Stadler, Sean P. Lane, M. Joy McClure,
Grace L. Jackson, Frederick D. Clavél, Masumi Iida, Marci E. J. Gleason,
Joy H. Xu, and Niall Bolger. 2018. {``Initial Elevation Bias in
Subjective Reports.''} \emph{PNAS} 115 (1): E15--23.
\url{https://doi.org/10.1073/pnas.1712277115}.

\leavevmode\vadjust pre{\hypertarget{ref-sullivan2017}{}}%
Sullivan, Edith V., Ty Brumback, Susan F. Tapert, Devin Prouty, Rosemary
Fama, Wesley K. Thompson, Sandra A. Brown, Kevin Cummins, Ian M.
Colrain, and Fiona C. Baker. 2017. {``Effects of Prior Testing Lasting a
Full Year in {NCANDA} Adolescents: Contributions from Age, Sex,
Socioeconomic Status, Ethnicity, Site, Family History of Alcohol or Drug
Abuse, and Baseline Performance.''} \emph{Developmental Cognitive
Neuroscience} 24: 72--83.

\leavevmode\vadjust pre{\hypertarget{ref-thompson2019}{}}%
Thompson, Wesley K., Deanna M. Barch, James M. Bjork, Raul Gonzalez,
Bonnie J. Nagel, Sara Jo Nixon, and Monica Luciana. 2019. {``The
Structure of Cognition in 9 and 10 Year-Old Children and Associations
with Problem Behaviors: {Findings} from the {ABCD} Study's Baseline
Neurocognitive Battery.''} \emph{Developmental Cognitive Neuroscience}
36: 100606.

\leavevmode\vadjust pre{\hypertarget{ref-uban2018}{}}%
Uban, Kristina A., Megan K. Horton, Joanna Jacobus, Charles Heyser,
Wesley K. Thompson, Susan F. Tapert, Pamela A. F. Madden, and Elizabeth
R. Sowell. 2018. {``Biospecimens and the {ABCD} Study: {Rationale},
Methods of Collection, Measurement and Early Data.''}
\emph{Developmental Cognitive Neuroscience}, The {Adolescent} {Brain}
{Cognitive} {Development} ({ABCD}) {Consortium}: {Rationale}, {Aims},
and {Assessment} {Strategy}, 32 (August): 97--106.
\url{https://doi.org/10.1016/j.dcn.2018.03.005}.

\leavevmode\vadjust pre{\hypertarget{ref-volkow2018}{}}%
Volkow, Nora D., George F. Koob, Robert T. Croyle, Diana W. Bianchi,
Joshua A. Gordon, Walter J. Koroshetz, Eliseo J. Pérez-Stable, et al.
2018. {``The Conception of the {ABCD} Study: {From} Substance Use to a
Broad {NIH} Collaboration.''} \emph{Developmental Cognitive
Neuroscience}, The {Adolescent} {Brain} {Cognitive} {Development}
({ABCD}) {Consortium}: {Rationale}, {Aims}, and {Assessment} {Strategy},
32 (August): 4--7. \url{https://doi.org/10.1016/j.dcn.2017.10.002}.

\leavevmode\vadjust pre{\hypertarget{ref-wang2019a}{}}%
Wang, Lijuan, Qian Zhang, Scott E. Maxwell, and C. S. Bergeman. 2019.
{``On Standardizing Within-Person Effects: {Potential} Problems of
Global Standardization.''} \emph{Multivariate Behavioral Research} 54
(3): 382--403.

\leavevmode\vadjust pre{\hypertarget{ref-watson2018}{}}%
Watson, Nicole, Eva Leissou, Heidi Guyer, and Mark Wooden. 2018. {``Best
{Practices} for {Panel} {Maintenance} and {Retention}.''} In
\emph{Advances in {Comparative} {Survey} {Methods}}, 597--622. John
Wiley \& Sons, Ltd. \url{https://doi.org/10.1002/9781118884997.ch29}.

\leavevmode\vadjust pre{\hypertarget{ref-zucker2018}{}}%
Zucker, Robert A., Raul Gonzalez, Sarah W. Feldstein Ewing, Martin P.
Paulus, Judith Arroyo, Andrew Fuligni, Amanda Sheffield Morris, Mariana
Sanchez, and Thomas Wills. 2018. {``Assessment of Culture and
Environment in the {Adolescent} {Brain} and {Cognitive} {Development}
{Study}: {Rationale}, Description of Measures, and Early Data.''}
\emph{Developmental Cognitive Neuroscience} 32: 107--20.

\end{CSLReferences}

\end{document}