### 2023-04-22 TailwindTest.qmd
added link to pdf version of manuscript

---
title: "Tailwind Test Doc"
editor: visual
---

ssss

<script src="https://cdn.tailwindcss.com"></script>

sssss

xxxx

:::{.text-xl .text-blue .rounded-lg .py-4 .px-2 .bg-gradient--to-r .from-green-700 .to-blue-900 .shadow-xl}

Place Content here

:::

xxxxx

.bg-gradient--to-r

.from-green-700

.to-blue-900

.shadow-xl

![](images/image-1130704809.png)