### 2023-04-22 TailwindTest.qmd
testing tailwind cdn

---
title: "Tailwind Test Doc"
editor: visual

format:
  html:
    css: styles.css

---
<!-- tailwind cdn temp link -->
<script src = "https://cdn.tailwindcss.com"></script>
<!---->

<div class="bg-gradient-to-r from-indigo-500 from-10% via-purple-500 via-30% to-pink-500 to-90% ...">
  <!-- ... -->
  xxxxxx
</div>

xxxxxx

<div class="bg-gradient-to-r from-indigo-500 from-10% via-purple-500 via-30% to-pink-500 to-90% ...">
  <!-- ... -->
  include content
</div>

:::{.text-xl .text-blue .rounded-lg .py-4 .px-2 .bg-gradient--to-r .from-green-700 .to-blue-900 .shadow-xl}

Place Content here

:::

<p .bg-sky-900>

include content

</p>

xxxxx

.bg-sky-900

.bg-gradient--to-r

.from-green-700

.to-blue-900

.shadow-xl