### 2023-04-23 TailwindTest.qmd
landing page update

---
title: "Tailwind Test Doc"
editor: visual

format:
  html:
    css: styles.css

---
<!-- tailwind cdn temp link -->
<script src = "https://cdn.tailwindcss.com"></script>
<!---->

<div class="bg-gradient-to-r from-indigo-500 from-10% via-purple-500 via-30% to-pink-500 to-90% ...">
  <!-- ... -->
  xxxzzzzzxxx
</div>

<div class="bg-gradient-to-r from-indigo-500 from-10% via-purple-500 via-30% to-pink-500 to-90% ...">
  <!-- ... -->
  include contentzzzzzzz
</div>

<div class="bg-gray-900">
  <!-- ... -->
  include content
</div>