### 2023-04-23 TailwindTest2.qmd

---
title: "TailwindTest2"
format:
  html:
    css: styles.css
    theme: cosmo
    fontsize: 1.1em
    linestretch: 1.7
---

<!-- tailwind cdn temp link -->
<script src = "https://cdn.tailwindcss.com"></script>
<!---->

<div class="bg-gradient-to-r from-indigo-500 from-10% via-purple-500 via-30% to-pink-500 to-90% ...">
This is github pages test site for the longitudinal analysis project. The information included within the site is for testing purposes only and contents may be inaccurate.
</div>

::: {.grid}

::: {.g-col-4}
This column takes 1/3 of the page
:::

::: {.g-col-8}
This column takes 2/3 of the page
![lights](./figures/lights.png)
:::

:::

zzzzzzzzz

---
#title: "Temporary Landing Page for Longitudinal Project"
# format: 
#   html:
#     page-layout: full
#     theme:
#       light: cosmos #flatly
#       dark: darkly
#     css: styles.css
#     smooth-scroll: true
#     linestretch: 1.7

---
<!-- tailwind cdn temp link -->
<script src = "https://cdn.tailwindcss.com"></script>
<!---->

<div class="bg-slate-400">

<div class="text-3xl text-center bg-gradient-to-r from-blue-700 from-10% via-blue-500 via-30% to-blue-300 to-90%">
This is the temporary landing page for the Longitudinal Analysis Project. <br> The site is currently under construction and used for testing purposes only. <br> Contents may be inaccurate.
</div>

<br>

::: {.grid}

::: {.g-col-2}
:::

::: {.g-col-8}
![](./figures/lights.png)
:::

::: {.g-col-2}
:::

:::

</div>