### 2023-04-23 ListofReferences.qmd

---
title: "List of Reference Material"
---

GLM @sec-glm
Non-independence @Non-independence

## GLM (Generalized Linear Model): 
A flexible generalization of ordinary linear regression that allows for response variables that have error distribution models other than a normal distribution.{#sec-glm}

## Non-independence: 
The situation where observations are not statistically independent, often arising in longitudinal or clustered data.

3.	Sample average of differences: The mean of the differences between paired observations, typically used in paired-sample designs to assess the average change between two time points or conditions.

4.	Fixed effects: Effects in a statistical model that are constant across all observations, representing the average relationship between predictors and the outcome variable.

5.	Random effects: Effects in a statistical model that vary across different levels of a grouping factor (e.g., subjects, clusters), capturing the variability within each level.

6.	Hierarchical data structures: Data structures where observations are nested within higher-order units (e.g., students within schools, employees within companies), resulting in dependency among observations.

7.	Within-subject correlations: Correlations between repeated measurements on the same subject or unit, often present in longitudinal or clustered data.

8.	Model fit indices: Quantitative measures used to assess how well a statistical model fits the data, such as R-squared, AIC, or BIC.

9.	Nonparametric: Statistical methods that do not rely on assumptions about the underlying distribution of the data.

10.	Resampling permutation method: A nonparametric method for hypothesis testing that involves randomly rearranging the observed data to create new samples and assess the probability of the observed test statistic under the null hypothesis.

11.	Clustered data: Data where observations are grouped or clustered within higher-order units, resulting in dependency among observations within each cluster.

12.	Population-averaged effects: Effects in a statistical model that represent the average relationship between predictors and the outcome variable across the entire population, rather than for specific levels of a grouping factor.

13.	Identity link: A link function in GLMs that relates the linear predictor to the expected value of the outcome variable without any transformation (E[Y] = Xβ).

14.	Logit link: A link function in GLMs used for binary outcomes, relating the linear predictor to the log-odds of the outcome variable (log(E[Y]/(1-E[Y])) = Xβ).

15.	Log link: A link function in GLMs used for count outcomes, relating the linear predictor to the natural logarithm of the expected value of the outcome variable (log(E[Y]) = Xβ).

16.	Logistic link: Synonymous with the logit link.

17.	Binomial distribution: A probability distribution that describes the number of successes in a fixed number of Bernoulli trials with the same probability of success.

18.	Exchangeable correlation structure: A working correlation structure in GEE models, assuming constant correlation between any pair of repeated measurements within subjects.

19.	Autoregressive correlation structure: A working correlation structure in GEE models, assuming correlation between repeated measurements decreases as the time lag between measurements increases.

20.	Unstructured correlation structure: A working correlation structure in GEE models, allowing for different correlations between each pair of repeated measurements within subjects.

21.	Sandwich estimator: A method for computing robust standard errors in GEE models, providing valid inference even if the working correlation structure is misspecified.

22.	Quasi-likelihood under the Independence model Criterion (QIC): A model selection criterion for GEE models, analogous to AIC, used to compare models with different working correlation structures.

23.	Odds ratios: A measure of association for binary outcomes, representing the ratio of the odds of an event occurring in one group to the odds in another group.

24.	Rate ratios: A measure of association for count outcomes, representing the ratio of the rates of an event occurring in one group to the rates in another group.

25.	Maximum likelihood estimation: A statistical method for estimating model parameters by finding the values that maximize the likelihood of the observed data given the model.

26.	Unbalanced data: Data where the number of observations or the number of measurements per subject or group is unequal across subjects or groups.

27.	Unequally spaced data: Data where the time intervals between repeated measurements or the spacing of observations are not equal.

28.	Maximum likelihood (ML): A method of estimation that involves finding the parameter values that maximize the likelihood of the observed data given a statistical model.

29.	Restricted maximum likelihood (REML): A modification of maximum likelihood estimation that is less biased for estimating variance components in mixed-effects models, particularly when the sample size is small.

30.	Penalized quasi-likelihood (PQL): An approximate method for fitting generalized linear mixed models, particularly for non-normal outcomes, which involves a penalized likelihood function and iteratively weighted least squares.

31.	Akaike Information Criterion (AIC): A model selection criterion that balances goodness-of-fit and model complexity, calculated as AIC = -2 * log-likelihood + 2 * k, where k is the number of parameters. Lower AIC values indicate better model fit.

32.	Bayesian Information Criterion (BIC): A model selection criterion that balances goodness-of-fit and model complexity, calculated as BIC = -2 * log-likelihood + k * log(n), where k is the number of parameters and n is the sample size. Lower BIC values indicate better model fit.