### 2023-04-24 howto.qmd

---
title: "How-To's: A Brief Overview"
---

The "how-to" section focuses on providing practical solutions to specific methods or problems users might encounter. Unlike tutorials, which aim to teach a concept, how-to guides concentrate on the efficient execution of specicif analyses without going into in-depth background explanations. By following these principles, the how-to section aims to create practical, efficient, and user-friendly guides that help users quickly identify solutions and accomplish their goals.

The how-to section should:

1. Be problem-oriented: How-to guides should address specific problems or analyses users may face, offering clear and concise solutions. Each guide should focus on a single issue, making it easy for users to find and apply the relevant information.

2. Offer step-by-step instructions: Similar to tutorials, how-to guides should provide a clear sequence of steps that users can follow to solve their problem or complete their analysis. Each step of the instructions should be explained in a straightforward manner.

3. Be concise: How-to guides should be to the point, offering just enough information to help users accomplish their goal. This means avoiding unnecessary explanations or background information, as users are typically looking for a quick solution.

4. Be actionable: The content in the how-to section should be practical, with a focus on actionable advice that users can apply immediately. This may involve providing code snippets, configuration settings, or other specific instructions that users can implement in their projects.

5. Include examples: How-to guides should include relevant examples that demonstrate the solution in action, allowing users to see how the instructions can be applied in a real-world context.

6. Cross-reference other resources: How-to guides should link to other relevant documentation, such as explanations or reference material, for users who may require additional background information or a deeper understanding of the topic.

For additional information/resources, see [https://documentation.divio.com/]