### 2023-04-25 ListofReferences.qmd
css updates

---
title: "List of Reference Material"
---

## Generalized Linear Model (GLM): 
A flexible generalization of ordinary linear regression that allows for response variables that have error distribution models other than a normal distribution.{#sec-glm}

## Non-independence: 
The situation where observations are not statistically independent, often arising in longitudinal or clustered data.

## Sample average of differences: 
The mean of the differences between paired observations, typically used in paired-sample designs to assess the average change between two time points or conditions.

## Fixed effects: 
Effects in a statistical model that are constant across all observations, representing the average relationship between predictors and the outcome variable.

## Random effects: 
Effects in a statistical model that vary across different levels of a grouping factor (e.g., subjects, clusters), capturing the variability within each level.

## Hierarchical data structures: 
Data structures where observations are nested within higher-order units (e.g., students within schools, employees within companies), resulting in dependency among observations.

## Within-subject correlations: 
Correlations between repeated measurements on the same subject or unit, often present in longitudinal or clustered data.

## Model fit indices: 
Quantitative measures used to assess how well a statistical model fits the data, such as R-squared, AIC, or BIC.

## Nonparametric: 
Statistical methods that do not rely on assumptions about the underlying distribution of the data.

## Resampling permutation method: 
A nonparametric method for hypothesis testing that involves randomly rearranging the observed data to create new samples and assess the probability of the observed test statistic under the null hypothesis.

## Clustered data: 
Data where observations are grouped or clustered within higher-order units, resulting in dependency among observations within each cluster.

## Population-averaged effects: 
Effects in a statistical model that represent the average relationship between predictors and the outcome variable across the entire population, rather than for specific levels of a grouping factor.

## Identity link: 
A link function in GLMs that relates the linear predictor to the expected value of the outcome variable without any transformation (E[Y] = Xβ).

## Logit link: 
A link function in GLMs used for binary outcomes, relating the linear predictor to the log-odds of the outcome variable (log(E[Y]/(1-E[Y])) = Xβ).

## Log link: 
A link function in GLMs used for count outcomes, relating the linear predictor to the natural logarithm of the expected value of the outcome variable (log(E[Y]) = Xβ).

## Logistic link: 
Synonymous with the logit link.

## Binomial distribution: 
A probability distribution that describes the number of successes in a fixed number of Bernoulli trials with the same probability of success.

## Exchangeable correlation structure: 
A working correlation structure in GEE models, assuming constant correlation between any pair of repeated measurements within subjects.

## Autoregressive correlation structure: 
A working correlation structure in GEE models, assuming correlation between repeated measurements decreases as the time lag between measurements increases.

## Unstructured correlation structure: 
A working correlation structure in GEE models, allowing for different correlations between each pair of repeated measurements within subjects.

## Sandwich estimator: 
A method for computing robust standard errors in GEE models, providing valid inference even if the working correlation structure is misspecified.

## Quasi-likelihood under the Independence model Criterion (QIC): 
A model selection criterion for GEE models, analogous to AIC, used to compare models with different working correlation structures.

## Odds ratios: 
A measure of association for binary outcomes, representing the ratio of the odds of an event occurring in one group to the odds in another group.

## Rate ratios: 
A measure of association for count outcomes, representing the ratio of the rates of an event occurring in one group to the rates in another group.

## Maximum likelihood estimation: 
A statistical method for estimating model parameters by finding the values that maximize the likelihood of the observed data given the model.

## Unbalanced data: 
Data where the number of observations or the number of measurements per subject or group is unequal across subjects or groups.

## Unequally spaced data: 
Data where the time intervals between repeated measurements or the spacing of observations are not equal.

## Maximum likelihood (ML): 
A method of estimation that involves finding the parameter values that maximize the likelihood of the observed data given a statistical model.

## Restricted maximum likelihood (REML): 
A modification of maximum likelihood estimation that is less biased for estimating variance components in mixed-effects models, particularly when the sample size is small.

## Penalized quasi-likelihood (PQL): 
An approximate method for fitting generalized linear mixed models, particularly for non-normal outcomes, which involves a penalized likelihood function and iteratively weighted least squares.

## Akaike Information Criterion (AIC): 
A model selection criterion that balances goodness-of-fit and model complexity, calculated as AIC = -2 * log-likelihood + 2 * k, where k is the number of parameters. Lower AIC values indicate better model fit.

## Bayesian Information Criterion (BIC): 
A model selection criterion that balances goodness-of-fit and model complexity, calculated as BIC = -2 * log-likelihood + k * log(n), where k is the number of parameters and n is the sample size. Lower BIC values indicate better model fit.