### 2023-04-25 explanations.qmd

---
title: "Explanations: A Brief Overview"
number-sections: true
format:
 html:
   css: styles.css
   theme: cosmo
   fontsize: 1.1em
   linestretch: 1.5
---

## What are Explanations?

::: {style="text-align: justify; font-size: 1.2em; line-stretch: 1.o; font-family: Georgia, serif;"}
[**The primary goal**]{style="color: goldenrod; font-size: 1.2em; font-family: Verdana, sans-serif;"} 
of the explanation section is to provide context, background information, and a deeper understanding of the concepts, methodologies, or processes related to a specific analysis or topic. Unlike tutorials and how-to guides, which are goal-oriented and problem-oriented respectively, explanations aim to clarify and enhance the user's knowledge. The explanation section aims to create informative, engaging, and accessible content that helps users build a solid foundation of understanding and knowledge about the subject matter.
:::

::: {style="text-align: justify; line-height: normal; color: black; border: solid black; background-color: antiquewhite; padding: 70px; font-family: Georgia, serif;"}

The explanation section should:

1. **Be informative**: Explanations should offer valuable insights and information, helping users develop a deeper understanding of the topic. This includes discussing the rationale behind certain choices, exploring alternative approaches, and explaining the advantages and disadvantages of various options.

2. **Address "why" questions**: Explanations should aim to answer the "why" questions users may have, delving into the reasoning, motivations, or principles that underlie a specific concept or approach.

3. **Be clear and concise**: While explanations may go into greater depth than tutorials or how-to guides, they should still be presented in a clear and concise manner. This means using plain language, avoiding jargon, and providing examples to help users grasp the concepts more easily.

4. **Be well-structured**: Explanations should be organized in a logical and coherent manner, making it easy for users to follow the flow of information and understand the relationships between different concepts.

5. **Support other documentation types**: Explanations should complement and support other types of documentation, such as tutorials and how-to guides, by providing the necessary context and background information. This may involve cross-referencing other resources or linking to related documentation.

6. **Engage the audience**: Explanations should be engaging and interesting, capturing the user's attention and encouraging them to explore the topic further. This may involve using analogies, storytelling, or real-world examples to illustrate the concepts and make them more relatable.

::: {style="text-align: center"}
For additional information, see [https://documentation.divio.com/]
:::

:::