### 2023-04-25 howto.qmd

---
title: "How-To's: A Brief Overview"
number-sections: true
format:
 html:
   css: styles.css
   theme: cosmo
   fontsize: 1.1em
   linestretch: 1.5
---

## What are How-To's?
::: {style="text-align: justify; font-size: 1.2em; line-stretch: 1.o; font-family: Georgia, serif;"}
[**The primary goal**]{style="color: goldenrod; font-size: 1.2em; font-family: Verdana, sans-serif;"} 
of the "how-to" section is on providing practical solutions to specific methods or analytic problems users might encounter. Unlike tutorials, which aim to teach a concept, how-to guides concentrate on the efficient execution of specific analyses without going into in-depth background explanations. By following these principles, the how-to section aims to create practical, efficient, and user-friendly guides that help users quickly identify solutions and accomplish their goals.
:::

::: {style="text-align: justify; line-height: normal; color: black; border: solid black; background-color: antiquewhite; padding: 70px; font-family: Georgia, serif;"}

The how-to section should:

1. **Be problem-oriented**: How-to guides should address specific problems or analyses users may face, offering clear and concise solutions. Each guide should focus on a single issue, making it easy for users to find and apply the relevant information.

2. **Offer step-by-step instructions**: Similar to tutorials, how-to guides should provide a clear sequence of steps that users can follow to solve their problem or complete their analysis. Each step of the instructions should be explained in a straightforward manner.

3. **Be concise**: How-to guides should be to the point, offering just enough information to help users accomplish their goal. This means avoiding unnecessary explanations or background information, as users are typically looking for a quick solution.

4. **Be actionable**: The content in the how-to section should be practical, with a focus on actionable advice that users can apply immediately. This may involve providing code snippets, configuration settings, or other specific instructions that users can implement in their projects.

5. **Include examples**: How-to guides should include relevant examples that demonstrate the solution in action, allowing users to see how the instructions can be applied in a real-world context.

6. **Cross-reference other resources**: How-to guides should link to other relevant documentation, such as explanations or reference material, for users who may require additional background information or a deeper understanding of the topic.

::: {style="text-align: center"}
For additional information, see [https://documentation.divio.com/]
:::

:::