### 2023-04-25 temp.qmd

---
title: "Tutorials: A Brief Overview"
number-sections: true
format:
 html:
   css: styles.css
   theme: cosmo
   fontsize: 1.1em
   linestretch: 1.7
---

This is the TUTORIAL main page example.

## Main Section

::: {style="color: red; border: solid black; background-color: grey;"}
This is the [middle section]{style="background-color: yellow; color: blue;"} 
:::

::: {.grid}

::: {.g-col-2}
:::

::: {.g-col-8}

::: {style="color: red; border: solid black; background-color: green;"}
Testing [Sample Test!]{style="background-color: yellow; color: blue;"}
:::

:::

::: {.g-col-2}
:::

:::

The primary goal of the tutorial section is to help users learn and understand a specific topic or method by providing an introductory and hands-on, step-by-step guide. This section aims to create engaging, accessible, and effective entry learning experiences that enable users to develop their skills and understanding of the subject matter.

<br>

::: {.grid}

::: {.g-col-2}
:::

::: {.g-col-8}
![](./figures/lights.png)
:::

::: {.g-col-2}
:::

:::