### 2023-04-26 temp.qmd

---
title: "Tutorials: A Brief Overview"
number-sections: true
format:
 html:
   css: styles.css
   theme: cosmo
   fontsize: 1.1em
   linestretch: 1.7
---

This is the TUTORIAL main page example.

## Main Section

```{ojs}
sunburst = {
  const root = partition(flareData);

  root.each(d => d.current = d);

  const svg = d3.create("svg")
      .attr("viewBox", [0, 0, width, width])
      .style("font", "15px sans-serif");

  const g = svg.append("g")
      .attr("transform", `translate(${width / 2},${width / 2})`);

  const path = g.append("g")
    .selectAll("path")
    .data(root.descendants().slice(1))
    .join("path")
      .attr("fill", d => { while (d.depth > 1) d = d.parent; return color(d.data.name); })
      .attr("fill-opacity", d => arcVisible(d.current) ? (d.children ? 0.6 : 0.4) : 0)
      .attr("d", d => arc(d.current));

  path.filter(d => d.children)
      .style("cursor", "pointer")
      .on("click", clicked);

  path.append("title")
      .text(d => `${d.ancestors().map(d => d.data.name).reverse().join("/")}\n${format(d.value)}`);

  const label = g.append("g")
      .attr("pointer-events", "none")
      .attr("text-anchor", "middle")
      .style("user-select", "none")
    .selectAll("text")
    .data(root.descendants().slice(1))
    .join("text")
      .attr("dy", "0.35em")
      .attr("fill-opacity", d => +labelVisible(d.current))
      .attr("transform", d => labelTransform(d.current))
      .text(d => d.data.name);

  const parent = g.append("circle")
      .datum(root)
      .attr("r", radius)
      .attr("fill", "none")
      .attr("pointer-events", "all")
      .on("click", clicked);

  function clicked(event, p) {
    parent.datum(p.parent || root);

    root.each(d => d.target = {
      x0: Math.max(0, Math.min(1, (d.x0 - p.x0) / (p.x1 - p.x0))) * 2 * Math.PI,
      x1: Math.max(0, Math.min(1, (d.x1 - p.x0) / (p.x1 - p.x0))) * 2 * Math.PI,
      y0: Math.max(0, d.y0 - p.depth),
      y1: Math.max(0, d.y1 - p.depth)
    });

    const t = g.transition().duration(750);

    // Transition the data on all arcs, even the ones that aren’t visible,
    // so that if this transition is interrupted, entering arcs will start
    // the next transition from the desired position.
    path.transition(t)
        .tween("data", d => {
          const i = d3.interpolate(d.current, d.target);
          return t => d.current = i(t);
        })
      .filter(function(d) {
        return +this.getAttribute("fill-opacity") || arcVisible(d.target);
      })
        .attr("fill-opacity", d => arcVisible(d.target) ? (d.children ? 0.6 : 0.4) : 0)
        .attrTween("d", d => () => arc(d.current));

    label.filter(function(d) {
        return +this.getAttribute("fill-opacity") || labelVisible(d.target);
      }).transition(t)
        .attr("fill-opacity", d => +labelVisible(d.target))
        .attrTween("transform", d => () => labelTransform(d.current));
  }

  function arcVisible(d) {
    return d.y1 <= 3 && d.y0 >= 1 && d.x1 > d.x0;
  }

  function labelVisible(d) {
    return d.y1 <= 3 && d.y0 >= 1 && (d.y1 - d.y0) * (d.x1 - d.x0) > 0.03;
  }

  function labelTransform(d) {
    const x = (d.x0 + d.x1) / 2 * 180 / Math.PI;
    const y = (d.y0 + d.y1) / 2 * radius;
    return `rotate(${x - 90}) translate(${y},0) rotate(${x < 180 ? 0 : 180})`;
  }

  return svg.node();
}
```

::: {style="color: red; border: solid black; background-color: grey;"}
This is the [middle section]{style="background-color: yellow; color: blue;"} 
:::

::: {.grid}

::: {.g-col-2}
:::

::: {.g-col-8}

::: {style="color: red; border: solid black; background-color: green;"}
Testing [Sample Test!]{style="background-color: yellow; color: blue;"}
:::

:::

::: {.g-col-2}
:::

:::

The primary goal of the tutorial section is to help users learn and understand a specific topic or method by providing an introductory and hands-on, step-by-step guide. This section aims to create engaging, accessible, and effective entry learning experiences that enable users to develop their skills and understanding of the subject matter.

<br>

::: {.grid}

::: {.g-col-2}
:::

::: {.g-col-8}
![](./figures/lights.png)
:::

::: {.g-col-2}
:::

:::