### 2023-04-26 documentation.qmd
update to documentation buttons

---
title: Documentation
author: Biostatistics Working Group
---

::: {.grid}
::: {.g-col-6}
::: {style="background-color: gray;"}
[**Tutorials**](./tutorials.qmd){.btn .btn-primary .btn-outline-warning .btn-lg .btn-block .btn role="link" data-toggle="tooltip" title="Link to the Tutorials documentation"}
:::
:::
::: {.g-col-6}
::: {style="background-color: gray;"}
[How-Tos](./howto.qmd){.btn .btn-primary  .btn-outline-warning .btn-lg .btn-block .btn role="button" data-toggle="tooltip" title="Link to the How-Tos documentation"}
:::
:::
:::

<br>

::: {.grid}
::: {.g-col-6}
::: {style="background-color: gray;"}
[Explanations](./explanations.qmd){.btn .btn-outline-warning .btn-lg .btn role="button"  data-toggle="tooltip" title="Link to the Explanations documentation"}
:::
:::
::: {.g-col-6}
::: {style="background-color: gray;"}
[Reference](./reference.qmd){.btn .btn-outline-warning .btn-lg .btn role="button" data-toggle="tooltip" title="Link to the Reference documentation"}
:::
:::
:::

<br>