### 2023-04-28 index.qmd
landing page updates 3

---
#title: "Longitudinal Analysis"
format:
  html:
    page-layout: full
    grid:
      sidebar-width: 0px
      body-width: 2000px
      margin-width: 0px
      gutter-width: 1.5rem
---

![](./figures/Figlanding6.svg)