### 2023-04-29 figures/Code for LDA Structure Mermaid Diagram.qmd
updated LDA Struct image for manuscript

---
title: "Code for LDA Structure Mermaid Diagram"
editor: visual
---

```{mermaid}

%%| echo: false
%%| fig-width: 3
%%| fig-height: 3

graph TB
A[LDA Structures] 
A -.-> |Focus on if & when an event occurs| B[Time-to-Event]
B -.-> C[Continuous-Time <br/>Survival Analysis]
B -.-> D[Discrete-Time <br/>Survival Analysis]
A -->|Focus on Change & Growth| E[Repeated Measures]
E -.-> F[Time-Series <br/> Models]
E -.-> G[Intensive <br/> Longitudinal Data]
E --> H[Longitudinal Panel Data]
H --Traditional Models--> J[Mean-level change & Marginal models]
H --Contemporary  Methods--> I[Within-person change & Random-effects models]

J --> K[ANOVA <br/> Generalized Estimating Equations <br/> ... ]
I --> L[Mixed Models <br/> Growth Curves <br/> ...]

click A href "https://www.github.com" "This is a link"
click B href "https://www.github.com" "This is a link"
click C href "https://www.github.com" "This is a link"
click D href "https://www.github.com" "This is a link"

style A fill:blue,stroke:#333,stroke-width:4px,color:#fff
style E fill:blue,stroke:#333,stroke-width:4px,color:#fff
style B fill:floralwhite,stroke:black
style C fill:floralwhite,stroke:black
style D fill:floralwhite,stroke:black
style F fill:floralwhite,stroke:black
style G fill:floralwhite,stroke:black

style H fill:blue,stroke:#333,stroke-width:4px,color:#fff
style J fill:blue,stroke:#333,stroke-width:2px,color:#fff
style I fill:blue,stroke:#333,stroke-width:2px,color:#fff
style K fill:blue,stroke:#333,stroke-width:2px,color:#fff,text-align:center
style L fill:blue,stroke:#333,stroke-width:2px,color:#fff,text-align:center

```