### 2023-04-30 index.qmd
updates to manucript figs

---
#title: "Longitudinal Analysis"
format:
  html:
    css: styles.css
    theme: cosmo
    page-layout: full
    grid:
      sidebar-width: 0px
      body-width: 2000px
      margin-width: 0px
    #gutter-width: 1.5rem
---

::: {.column-screen}
![](./figures/Figlanding5.svg)
:::