### 2023-04-30 temp.qmd

---
#title: "Tutorials: A Brief Overview"
format:
  html:
    page-layout: full
    grid:
      #sidebar-width: 250px
      sidebar-width: 0px
      body-width: 2000px
      margin-width: 0px
      #gutter-width: 1.5rem
    #mermaid:
      #theme: forest
    css: styles.css
#    theme: cosmo
#    fontsize: 1.1em
#    linestretch: 1.7

---

<script src="https://cdn.tailwindcss.com"></script>

```{html}
<div class="alert alert-primary" role="alert">
  This is a primary alert—check it out!
</div>
```

xxxyyyzzz
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```       
\begin{tabular}[t]{@{}l}
This is where the reciepient goes \\ 
Address 1 \\
Address 2 \\
Address 3 \\
Address 3 \\
\end{tabular}
\hfill
\begin{tabular}[t]{@{}l}
This is where the sender goes \\  
Person \\
Department \\ 
Address 1 \\ 
Address 2 \\  
Address 3 \\  
Address 4 \\  
Address 5 \\  
Address 6 \\ 
Telephone \\  
Fax \\  
Email \\
\end{tabular}

xxxyyyzzz
xxxxx

<div class="alert alert-primary" role="alert">
  This is a primary alert with <a href="#" class="alert-link">an example link</a>. Give it a click if you like.
</div>

<span class="badge badge-pill badge-primary">Primary</span>

<a href="#" class="badge badge-primary">Primary</a>

<div class="card" style="width: 18rem;">
  <img class="card-img-top" src="..." alt="Card image cap">
  <div class="card-body">
    <div class="card-title">Card title</div>
    <p class="card-text">Some quick example text to build on the card title and make up the bulk of the card's content.</p>
    <a href="#" class="btn btn-primary">Go somewhere</a>
  </div>
</div>

<div class="card">
  <div class="card-body">
    This is some text within a card body.
  </div>
</div>

<div class="card" style="width: 18rem;">
  <img class="card-img-top" src="..." alt="Card image cap">
  <div class="card-body">
    <h5 class="card-title">Card title</h5>
    <p class="card-text">Some quick example text to build on the card title and make up the bulk of the card's content.</p>
  </div>
  <ul class="list-group list-group-flush">
    <li class="list-group-item">Cras justo odio</li>
    <li class="list-group-item">Dapibus ac facilisis in</li>
    <li class="list-group-item">Vestibulum at eros</li>
  </ul>
  <div class="card-body">
    <a href="#" class="card-link">Card link</a>
    <a href="#" class="card-link">Another link</a>
  </div>
</div>

<div class="card">
  <div class="card-header">
    Featured
  </div>
  <div class="card-body">
    <h5 class="card-title">Special title treatment</h5>
    <p class="card-text">With supporting text below as a natural lead-in to additional content.</p>
    <a href="#" class="btn btn-primary">Go somewhere</a>
  </div>
</div>

<div class="card">
  <h5 class="card-header">Featured</h5>
  <div class="card-body">
    <h5 class="card-title">Special title treatment</h5>
    <p class="card-text">With supporting text below as a natural lead-in to additional content.</p>
    <a href="#" class="btn btn-primary">Go somewhere</a>
  </div>
</div>

<div class="card text-white bg-primary mb-3" style="max-width: 18rem;">
  <div class="card-header">Header</div>
  <div class="card-body">
    <h5 class="card-title">Primary card title</h5>
    <p class="card-text">Some quick example text to build on the card title and make up the bulk of the card's content.</p>
  </div>
</div>

<div class="card border-primary mb-3" style="max-width: 18rem;">
  <div class="card-header">Header</div>
  <div class="card-body text-primary">
    <h5 class="card-title">Primary card title</h5>
    <p class="card-text">Some quick example text to build on the card title and make up the bulk of the card's content.</p>
  </div>
</div>

<div class="card border-success mb-3" style="max-width: 18rem;">
  <div class="card-header bg-transparent border-success">Header</div>
  <div class="card-body text-success">
    <h5 class="card-title">Success card title</h5>
    <p class="card-text">Some quick example text to build on the card title and make up the bulk of the card's content.</p>
  </div>
  <div class="card-footer bg-transparent border-success">Footer</div>
</div>
<div class="alert alert-success" role="alert">
  <h4 class="alert-heading">Well done!</h4>
  <p>Aww yeah, you successfully read this important alert message. This example text is going to run a bit longer so that you can see how spacing within an alert works with this kind of content.</p>
  <hr>
  <p class="mb-0">Whenever you need to, be sure to use margin utilities to keep things nice and tidy.</p>
</div>

<dl class="row">
  <dt class="col-sm-3">Description lists</dt>
  <dd class="col-sm-9">A description list is perfect for defining terms.</dd>

  <dt class="col-sm-3">Euismod</dt>
  <dd class="col-sm-9">
    <p>Vestibulum id ligula porta felis euismod semper eget lacinia odio sem nec elit.</p>
    <p>Donec id elit non mi porta gravida at eget metus.</p>
  </dd>

  <dt class="col-sm-3">Malesuada porta</dt>
  <dd class="col-sm-9">Etiam porta sem malesuada magna mollis euismod.</dd>

  <dt class="col-sm-3 text-truncate">Truncated term is truncated</dt>
  <dd class="col-sm-9">Fusce dapibus, tellus ac cursus commodo, tortor mauris condimentum nibh, ut fermentum massa justo sit amet risus.</dd>

  <dt class="col-sm-3">Nesting</dt>
  <dd class="col-sm-9">
    <dl class="row">
      <dt class="col-sm-4">Nested definition list</dt>
      <dd class="col-sm-8">Aenean posuere, tortor sed cursus feugiat, nunc augue blandit nunc.</dd>
    </dl>
  </dd>
</dl>

<!-- Button trigger modal -->
<button type="button" class="btn btn-primary" data-toggle="modal" data-target="#exampleModal">
  Launch demo modal
</button>

<!-- Modal -->
<div class="modal fade" id="exampleModal" tabindex="-1" role="dialog" aria-labelledby="exampleModalLabel" aria-hidden="true">
  <div class="modal-dialog" role="document">
    <div class="modal-content">
      <div class="modal-header">
        <h5 class="modal-title" id="exampleModalLabel">Modal title</h5>
        <button type="button" class="close" data-dismiss="modal" aria-label="Close">
          <span aria-hidden="true">&times;</span>
        </button>
      </div>
      <div class="modal-body">
        ...
      </div>
      <div class="modal-footer">
        <button type="button" class="btn btn-secondary" data-dismiss="modal">Close</button>
        <button type="button" class="btn btn-primary">Save changes</button>
      </div>
    </div>
  </div>
</div>

<button type="button" class="btn btn-lg btn-danger" data-toggle="popover" title="Popover title" data-content="And here's some amazing content. It's very engaging. Right?">Click to toggle popover</button>

<div class="w-25 p-3" style="background-color: #eee;">Width 25%</div>
<div class="w-50 p-3" style="background-color: #eee;">Width 50%</div>
<div class="w-75 p-3" style="background-color: #eee;">Width 75%</div>
<div class="w-100 p-3" style="background-color: #eee;">Width 100%</div>

<div style="height: 100px; background-color: rgba(255,0,0,0.1);">
  <div class="h-25 d-inline-block" style="width: 120px; background-color: rgba(0,0,255,.1)">Height 25%</div>
  <div class="h-50 d-inline-block" style="width: 120px; background-color: rgba(0,0,255,.1)">Height 50%</div>
  <div class="h-75 d-inline-block" style="width: 120px; background-color: rgba(0,0,255,.1)">Height 75%</div>
  <div class="h-100 d-inline-block" style="width: 120px; background-color: rgba(0,0,255,.1)">Height 100%</div>
</div>

<p class="text-justify">Ambitioni dedisse scripsisse iudicaretur. Cras mattis iudicium purus sit amet fermentum. Donec sed odio operae, eu vulputate felis rhoncus. Praeterea iter est quasdam res quas ex communi. At nos hinc posthac, sitientis piros Afros. Petierunt uti sibi concilium totius Galliae in diem certam indicere. Cras mattis iudicium purus sit amet fermentum.</p>

<p class="text-left">Left aligned text on all viewport sizes.</p>
<p class="text-center">Center aligned text on all viewport sizes.</p>
<p class="text-right">Right aligned text on all viewport sizes.</p>

<p class="text-sm-left">Left aligned text on viewports sized SM (small) or wider.</p>
<p class="text-md-left">Left aligned text on viewports sized MD (medium) or wider.</p>
<p class="text-lg-left">Left aligned text on viewports sized LG (large) or wider.</p>
<p class="text-xl-left">Left aligned text on viewports sized XL (extra-large) or wider.</p>

<p class="font-weight-bold">Bold text.</p>
<p class="font-weight-normal">Normal weight text.</p>
<p class="font-weight-light">Light weight text.</p>
<p class="font-italic">Italic text.</p>

<table class="table table-bordered">
  <thead>
    <tr>
      <th scope="col">#</th>
      <th scope="col">First</th>
      <th scope="col">Last</th>
      <th scope="col">Handle</th>
    </tr>
  </thead>
  <tbody>
    <tr>
      <th scope="row">1</th>
      <td>Mark</td>
      <td>Otto</td>
      <td>@mdo</td>
    </tr>
    <tr>
      <th scope="row">2</th>
      <td>Jacob</td>
      <td>Thornton</td>
      <td>@fat</td>
    </tr>
    <tr>
      <th scope="row">3</th>
      <td colspan="2">Larry the Bird</td>
      <td>@twitter</td>
    </tr>
  </tbody>
</table>

<blockquote class="blockquote">
  <p class="mb-0">Lorem ipsum dolor sit amet, consectetur adipiscing elit. Integer posuere erat a ante.</p>
</blockquote>

<table class="table">
  <thead>
    <tr>
      <th scope="col">#</th>
      <th scope="col">First</th>
      <th scope="col">Last</th>
      <th scope="col">Handle</th>
    </tr>
  </thead>
  <tbody>
    <tr>
      <th scope="row">1</th>
      <td>Mark</td>
      <td>Otto</td>
      <td>@mdo</td>
    </tr>
    <tr>
      <th scope="row">2</th>
      <td>Jacob</td>
      <td>Thornton</td>
      <td>@fat</td>
    </tr>
    <tr>
      <th scope="row">3</th>
      <td>Larry</td>
      <td>the Bird</td>
      <td>@twitter</td>
    </tr>
  </tbody>
</table>

<table class="hover:table-fixed">
  <thead>
    <tr>
      <th>Song</th>
      <th>Artist</th>
      <th>Year</th>
    </tr>
  </thead>
  <tbody>
    <tr>
      <td>The Sliding Mr. Bones (Next Stop, Pottersville)</td>
      <td>Malcolm Lockyer</td>
      <td>1961</td>
    </tr>
    <tr>
      <td>Witchy Woman</td>
      <td>The Eagles</td>
      <td>1972</td>
    </tr>
    <tr>
      <td>Shining Star</td>
      <td>Earth, Wind, and Fire</td>
      <td>1975</td>
    </tr>
  </tbody>
</table>

::: {.column-screen}
![-](./figures/Figlanding5.svg)
:::

xxxxx

file:///Users/shawes/git/swhawes.github.io/docs/manuscript.html
xxxxx

```{dot}
digraph "Linux_kernel_diagram" {
	fontname="Helvetica,Arial,sans-serif"
	node [fontname="Helvetica,Arial,sans-serif"]
	edge [fontname="Helvetica,Arial,sans-serif"]
	graph [
		newrank = true,
		nodesep = 0.3,
		ranksep = 0.2,
		overlap = true,
		splines = false,
	]
	node [
		fixedsize = false,
		fontsize = 24,
		height = 1,
		shape = box,
		style = "filled,setlinewidth(5)",
		width = 2.2
	]
	edge [
		arrowhead = none,
		arrowsize = 0.5,
		labelfontname = "Ubuntu",
		weight = 10,
		style = "filled,setlinewidth(5)"
	]
	subgraph system {
		node [color = "#e27dd6ff"]
		edge [color = "#e27dd6ff"]
		system_ [
			fixedsize = true,
			height = 0,
			shape = point,
			style = invis,
			shape = point
		]
		system [
			URL = "https://en.wikibooks.org/wiki/The_Linux_Kernel/System",
			fillcolor = white,
			fixedsize = true,
			height = 0.6,
			row = func,
			width = 2]
		system -> system_ [
			arrowhead = "",
			row = func];
		SCI [
			URL = "https://en.wikibooks.org/wiki/The_Linux_Kernel/Syscalls",
			fillcolor = "#d9e7ee",
			fixedsize = true,
			label = "System calls",
			row = usr,
			shape = ellipse]
		sysfs [
			fillcolor = "#b2d3e4",
			label = "proc & sysfs\nfile systems"]
		SCI -> sysfs
		DM [
			fillcolor = "#91b5c9",
			fixedsize = true,
			fontsize = 20,
			height = 0.8,
			label = "Device\nModel",
			shape = octagon,
			width = 2]
		sysfs -> DM
		log_sys [
			fillcolor = "#6a9ab1",
			fontsize = 20,
			label = "system run,\nmodules,\ngeneric\nHW access "]
		DM -> log_sys
		bus_drv [
			fillcolor = "#71809b",
			label = "bus drivers"]
		log_sys -> bus_drv
		buses [
			fillcolor = "#777777",
			fontcolor = white,
			fontsize = 20,
			label = "buses:\nPCI, USB ...",
			row = chip]
		bus_drv -> buses
	}
	subgraph networking {
		node [color = "#61c2c5"]
		edge [color = "#61c2c5"]
		networking_ [
			fixedsize = true,
			height = 0,
			shape = point,
			style = invis,
			shape = point
				width = 0]
		networking [
			URL = "https://en.wikibooks.org/wiki/The_Linux_Kernel/Networking",
			fillcolor = white,
			fixedsize = true,
			height = 0.6,
			row = func,
			width = 2]
		networking -> networking_ [
			arrowhead = "",
			row = func]
		sock [
			fillcolor = "#d9e7ee",
			fixedsize = true,
			label = Sockets,
			row = usr,
			shape = ellipse]
		prot_fam [
			fillcolor = "#b2d3e4",
			label = "protocol\nfamilies"]
		sock -> prot_fam
		log_prot [
			fillcolor = "#6a9ab1",
			label = "protocols:\nTCP, UDP, IP"]
		prot_fam -> log_prot
		netif [
			fillcolor = "#71809b",
			fontsize = 20,
			label = "network\ninterfaces\nand drivers"]
		log_prot -> netif
		net_hw [
			fillcolor = "#777777",
			fontcolor = white,
			fontsize = 20,
			label = "network:\nEthernet, WiFi ...",
			row = chip]
		netif -> net_hw
		NFS [
			color = "#8383cc",
			fillcolor = "#91b5c9",
			fixedsize = true,
			height = 0.8,
			label = NFS,
			shape = octagon,
			width = 1.2]
		NFS -> log_prot [weight = 0]
	}
	subgraph processing {
		node [color = "#c46747"]
		edge [color = "#c46747"]
		processing_ [
			fixedsize = true,
			height = 0,
			shape = point
				style = invis,
			width = 0]
		processing [
			URL = "https://en.wikibooks.org/wiki/The_Linux_Kernel/Processing",
			fillcolor = white,
			fixedsize = true,
			height = 0.6,
			row = func,
			width = 2]
		processing -> processing_ [
			arrowhead = "",
			row = func]
		proc [
			fillcolor = "#d9e7ee",
			fixedsize = true,
			label = Processes,
			row = usr,
			shape = ellipse]
		Tasks [
			fillcolor = "#b2d3e4"]
		proc -> Tasks
		sync [
			fillcolor = "#91b5c9",
			fixedsize = true,
			fontsize = 20,
			fontname = "Arial Narrow"
			label = synchronization,
			height = 0.7,
			//width = 2,
			shape = octagon]
		Tasks -> sync
		sched [
			fillcolor = "#6a9ab1",
			label = Scheduler]
		sync -> sched
		IRQ [
			fillcolor = "#71809b",
			fontsize = 20,
			label = "interrupts\ncore,\nCPU arch"]
		sched -> IRQ
		CPU [
			fillcolor = "#777777",
			fontcolor = white,
			fontsize = 20,
			row = chip]
		IRQ -> CPU
	}	// processing
	subgraph mem {
		node [
			color = "#51bf5b",
			height = 1
		]
		edge [color = "#51bf5b"]
		MA [
			color = "#51bf5b",
			fillcolor = "#d9e7ee",
			fixedsize = true,
			label = "memory\naccess",
			row = usr,
			height = 1,
			shape = ellipse]
		MA -> VM
		mmap [
			fillcolor = "#91b5c9",
			fixedsize = true,
			fontsize = 20,
			height = 0.8,
			label = "memory\nmapping",
			shape = octagon,
			width = 2]
		mmap -> log_mem
		log_mem -> PA
		SW [
			color = "#8383cc",
			fillcolor = "#91b5c9",
			fixedsize = true,
			label = Swap,
			height = 0.8,
			shape = octagon,
			width = 1.2]
		mmap -> SW [weight = 1]
		SW -> block [
			color = "#8383cc", weight = 1]
		PA [
			fillcolor = "#71809b",
			label = "Page\nAllocator"
		]
		PC -> PA [weight = 0 color="#51bf5b"]
		RAM [
			color = "#51bf5b",
			fillcolor = "#777777",
			fontcolor = white,
			fontsize = 20,
			label = "MMU, RAM",
			height = 1,
			row = chip]
		PA -> RAM
		memory -> memory_ [
			arrowhead = "",
			row = func]
		VM -> mmap
	}	// mem
	subgraph storage {
		node [color = "#8383cc"]
		edge [color = "#8383cc"]
		NFS;
		storage_ [
			shape = point,
			fixedsize = true,
			height = 0,
			style = invis,
			width = 0]
		storage [
			URL = "https://en.wikibooks.org/wiki/The_Linux_Kernel/Storage",
			fillcolor = white,
			fixedsize = true,
			height = 0.6,
			row = func,
			width = 2]
		storage -> storage_ [
			arrowhead = "",
			row = func]
		FS [
			fillcolor = "#d9e7ee",
			fixedsize = true,
			label = "files and\ndirectories",
			row = usr,
			shape = ellipse]
		VFS [
			fillcolor = "#b2d3e4",
			label = "Virtual\nFile System"]
		FS -> VFS
		VFS -> mmap [weight = 0]
		VFS -> NFS [weight = 0]
		logFS [
			fillcolor = "#6a9ab1",
			fontsize = 20,
			label = "logical\nfilesystems:\next3, xfs ..."]
		VFS -> logFS
		PC [
			fillcolor = "#91b5c9",
			fixedsize = true,
			fontsize = 20,
			height = 0.8,
			label = "page\ncache",
			shape = octagon,
			width = 1.2]
		VFS -> PC [weight = 0]
		block [
			fillcolor = "#71809b",
			fontsize = 20,
			label = "Block\ndevices\nand drivers"]
		logFS -> block
		SD [
			fillcolor = "#777777",
			fontcolor = white,
			fontsize = 20,
			label = "storage devices:\nSCSI, NVMe ...",
			row = chip]
		block -> SD
	}	// storge
	subgraph HI {
		node [color = "#cfbf57ff"]
		edge [
			color = "#cfbf57ff",
			weight = 10
		]
		HI_ [
			fixedsize = true,
			height = 0,
			shape = point,
			style = invis,
			width = 0]
		HI [
			URL = "https://en.wikibooks.org/wiki/The_Linux_Kernel/Human_interfaces",
			fillcolor = white,
			fixedsize = true,
			fontsize = 12,
			height = 0.6,
			label = "Forms\nof\nchange",
			row = func,
			width = 2]
		HI -> HI_ [
			arrowhead = "",
			row = func]
		char [
			fillcolor = "#d9e7ee",
			fixedsize = true,
			label = "char\ndevices",
			row = usr,
			shape = ellipse]
		input [
			fillcolor = "#b2d3e4",
			label = "input\nsubsystem"]
		char -> input
		F7 [
			fillcolor = "#6a9ab1",
			label = "HI class\ndrivers"]
		input -> F7
		HID [
			fillcolor = "#71809b",
			fontsize = 20,
			URL = "https://www.kernel.org/doc/html/latest/hid/",
			label = "HI\nperipherals\ndrivers"]
		F7 -> HID
		display [
			fillcolor = "#777777",
			fontcolor = white,
			fontsize = 19,
			label = "keyboard, mouse,\ndisplay, audio",
			row = chip]
		HID -> display
	} // HI
	subgraph functions {
		graph [rank = same]
		edge [
			style = invis,
			weight = 1
		]
		system;
		networking;
		system -> processing [weight = 1]
		storage -> networking [weight = 1]
		memory [
			color = "#51bf5b",
			URL = "https://en.wikibooks.org/wiki/The_Linux_Kernel/Memory",
			fillcolor = white,
			fixedsize = true,
			height = 0.6,
			row = func,
			width = 2]
		memory -> storage [weight = 1]
		processing -> memory [weight = 1]
		functions_ [
			fixedsize = true,
			height = 0,
			shape = point
			style = invis,
			width = 0]
		functions_ -> HI -> system [weight = 1]
		functions [
			color = gray,
			tooltip = "Columns represent main functionalities of the kernel",
			URL = "http://www.makelinux.net/ldd3/chp-1-sect-2.shtml",
			fillcolor = gray,
			fixedsize = true,
			height = 0.6,
			row = func,
			style = dashed,
			width = 1.6]
		functions -> functions_ [
			arrowhead = "",
			color = gray,
			style = "",
			weight = ""]
	}
	subgraph interfaces {
		graph [rank = same]
		SCI;
		sock;
		FS;
		proc;
		char;
		usr_ [
			fixedsize = true,
			height = 0,
			shape = point
				style = invis,
			width = 0.5]
		usr [
			fillcolor = "#d9e7eeff",
			fixedsize = true,
			label = "user space\ninterfaces",
			row = usr,
			shape = ellipse,
			style = "filled,setlinewidth(0)"]
		MA;
	}
	{
		edge [style = invis weight = 10 ]
		system_;
		SCI;
		system_ -> SCI;
		networking_;
		sock;
		networking_ -> sock;
		storage_;
		FS;
		storage_ -> FS;
		processing_;
		proc;
		processing_ -> proc;
		HI_;
		char;
		HI_ -> char;
		MA;
		memory_ [
			fixedsize = true,
			height = 0,
			shape = point,
			style = invis,
			width = 0]
		memory_ -> MA;
	}
	subgraph virtual {
		graph [rank = same]
		sysfs;
		prot_fam;
		VFS;
		Tasks;
		input;
		D0 [
			fixedsize = true,
			height = 0,
			shape = point,
			style = invis,
			width = 0]
		virt [
			fillcolor = "#b2d3e4",
			label = "virtual\nsubsystems",
			URL = "https://en.wikipedia.org/wiki/Proxy_pattern",
			tooltip = "proxy between standard user space interfaces and internal implementations",
			style = "filled,setlinewidth(0)"]
		VM [
			color = "#51bf5b",
			fillcolor = "#b2d3e4",
			label = "Virtual\nmemory"]
	}
	subgraph bridges {
		graph [rank = same]
		bridges [
			fillcolor = "#91b5c9",
			shape = octagon,
			tooltip = "bridges between uniform virtual interfaces and various implementations",
			URL = "https://en.wikipedia.org/wiki/Bridge_pattern",
			style = "filled,setlinewidth(0)"]
		DM;
		NFS;
		mmap;
		sync;
		E0 [
			fixedsize = true,
			height = 0,
			shape = point,
			style = invis,
			width = 0]
		//PC
	}
	subgraph logical {
		graph [rank = same]
		log_sys;
		log_prot;
		logFS;
		sched;
		F7;
		F0 [
			fixedsize = true,
			height = 0,
			shape = point,
			style = invis,
			width = 0]
		logical [
			fillcolor = "#6a9ab1",
			style = "filled,setlinewidth(0)"]
		log_mem [
			color = "#51bf5b",
			fillcolor = "#6a9ab1",
			label = "logical\nmemory"]
		//SW
	}
	subgraph HWI {
		graph [rank = same]
		HWI [
			fillcolor = "#71809b",
			label = "hardware\ninterfaces",
			style = "filled,setlinewidth(0)"]
		bus_drv;
		netif;
		block;
		//PA;
		IRQ;
		HID;
		G0 [
			fixedsize = true,
			height = 0,
			shape = point,
			style = invis,
			width = 0]
	}
	subgraph HW {
		graph [rank = same]
		HW [
			fillcolor = "#777777",
			fontcolor = white,
			label = "electronics,\nhardware",
			row = chip,
			style = "filled,setlinewidth(0)"]
		buses;
		net_hw;
		SD;
		CPU;
		display;
		H0 [
			fixedsize = true,
			height = 0,
			shape = point,
			style = invis,
			width = 0]
		RAM;
	}
	bottom [
		label = "© 2007-2022 Costa Shulyupin http://www.MakeLinux.net/kernel/diagram",
		URL = "http://www.MakeLinux.net/kernel/diagram",
		shape = plaintext,
		style = ""]
	CPU -> bottom [style = invis]
	layers [
		fillcolor = lightgray,
		tooltip = "Functionalities are divided to common layers. It is approximate division.",
		height = 0.1,
		style = "filled,setlinewidth(0)",
		width = 0.5]
	functions -> layers [style = invis ]
	usr -> usr_ [
		arrowhead = "",
		color = "#d9e7eeff",
		minlen = 2]
	usr -> virt [
		color = "#d9e7eeff"]
	virt -> D0 [
		arrowhead = "",
		color = "#b2d3e4",
		minlen = 2]
	virt -> bridges [
		color = "#b2d3e4"]
	bridges -> E0 [
		arrowhead = "",
		color = "#91b5c9",
		minlen = 2,
		style = "filled,setlinewidth(6)",
		weight = ""]
	bridges -> logical [
		color = "#91b5c9",
		style = "filled,setlinewidth(6)"]
	logical -> F0 [
		arrowhead = "",
		color = "#6a9ab1",
		minlen = 2,
		row = logical,
		style = "filled,setlinewidth(6)",
		weight = ""]
	logical -> HWI [
		color = "#6a9ab1",
		row = logical,
		style = "filled,setlinewidth(6)"]
	HWI -> G0 [
		arrowhead = "",
		color = "#71809b",
		minlen = 2,
		row = HWI,
		style = "filled,setlinewidth(6)",
		weight = ""]
	HWI -> HW [
		color = "#71809b",
		row = HWI,
		style = "filled,setlinewidth(6)"]
	HW -> H0 [
		arrowhead = "",
		color = "#777777",
		minlen = 2,
		row = chip,
		style = "filled,setlinewidth(6)",
		weight = ""]
	layers -> usr [
		arrowhead = "",
		color = gray,
		style = "filled,setlinewidth(1)"]
	LKD [
		fontsize = 40,
		label = "Linux kernel diagram",
		shape = plain,
		style = ""]
	LKD -> processing [style = invis]
}

```

xxxxx

![LPA Structure](./figures/LDA Structure 3.svg)

## Section III

Section focused on different ways of considering/grouping/deciding on which specific analysis to choose (see curran&bauer slides)

```{dot}
digraph "Linux_kernel_diagram" {
	fontname="Helvetica,Arial,sans-serif"
	node [fontname="Helvetica,Arial,sans-serif"]
	edge [fontname="Helvetica,Arial,sans-serif"]
	graph [
		newrank = true,
		nodesep = 0.3,
		ranksep = 0.2,
		overlap = true,
		splines = false,
	]
	node [
		fixedsize = false,
		fontsize = 24,
		height = 1,
		shape = box,
		style = "filled,setlinewidth(5)",
		width = 2.2
	]
	edge [
		arrowhead = none,
		arrowsize = 0.5,
		labelfontname = "Ubuntu",
		weight = 10,
		style = "filled,setlinewidth(5)"
	]
	subgraph system {
		node [color = "#e27dd6ff"]
		edge [color = "#e27dd6ff"]
		system_ [
			fixedsize = true,
			height = 0,
			shape = point,
			style = invis,
			shape = point
		]
		system [
			URL = "https://en.wikibooks.org/wiki/The_Linux_Kernel/System",
			fillcolor = white,
			fixedsize = true,
			height = 0.6,
			row = func,
			width = 2]
		system -> system_ [
			arrowhead = "",
			row = func];
		SCI [
			URL = "https://en.wikibooks.org/wiki/The_Linux_Kernel/Syscalls",
			fillcolor = "#d9e7ee",
			fixedsize = true,
			label = "System calls",
			row = usr,
			shape = ellipse]
		sysfs [
			fillcolor = "#b2d3e4",
			label = "proc & sysfs\nfile systems"]
		SCI -> sysfs
		DM [
			fillcolor = "#91b5c9",
			fixedsize = true,
			fontsize = 20,
			height = 0.8,
			label = "Device\nModel",
			shape = octagon,
			width = 2]
		sysfs -> DM
		log_sys [
			fillcolor = "#6a9ab1",
			fontsize = 20,
			label = "system run,\nmodules,\ngeneric\nHW access "]
		DM -> log_sys
		bus_drv [
			fillcolor = "#71809b",
			label = "bus drivers"]
		log_sys -> bus_drv
		buses [
			fillcolor = "#777777",
			fontcolor = white,
			fontsize = 20,
			label = "buses:\nPCI, USB ...",
			row = chip]
		bus_drv -> buses
	}
	subgraph networking {
		node [color = "#61c2c5"]
		edge [color = "#61c2c5"]
		networking_ [
			fixedsize = true,
			height = 0,
			shape = point,
			style = invis,
			shape = point
				width = 0]
		networking [
			URL = "https://en.wikibooks.org/wiki/The_Linux_Kernel/Networking",
			fillcolor = white,
			fixedsize = true,
			height = 0.6,
			row = func,
			width = 2]
		networking -> networking_ [
			arrowhead = "",
			row = func]
		sock [
			fillcolor = "#d9e7ee",
			fixedsize = true,
			label = Sockets,
			row = usr,
			shape = ellipse]
		prot_fam [
			fillcolor = "#b2d3e4",
			label = "protocol\nfamilies"]
		sock -> prot_fam
		log_prot [
			fillcolor = "#6a9ab1",
			label = "protocols:\nTCP, UDP, IP"]
		prot_fam -> log_prot
		netif [
			fillcolor = "#71809b",
			fontsize = 20,
			label = "network\ninterfaces\nand drivers"]
		log_prot -> netif
		net_hw [
			fillcolor = "#777777",
			fontcolor = white,
			fontsize = 20,
			label = "network:\nEthernet, WiFi ...",
			row = chip]
		netif -> net_hw
		NFS [
			color = "#8383cc",
			fillcolor = "#91b5c9",
			fixedsize = true,
			height = 0.8,
			label = NFS,
			shape = octagon,
			width = 1.2]
		NFS -> log_prot [weight = 0]
	}
	subgraph processing {
		node [color = "#c46747"]
		edge [color = "#c46747"]
		processing_ [
			fixedsize = true,
			height = 0,
			shape = point
				style = invis,
			width = 0]
		processing [
			URL = "https://en.wikibooks.org/wiki/The_Linux_Kernel/Processing",
			fillcolor = white,
			fixedsize = true,
			height = 0.6,
			row = func,
			width = 2]
		processing -> processing_ [
			arrowhead = "",
			row = func]
		proc [
			fillcolor = "#d9e7ee",
			fixedsize = true,
			label = Processes,
			row = usr,
			shape = ellipse]
		Tasks [
			fillcolor = "#b2d3e4"]
		proc -> Tasks
		sync [
			fillcolor = "#91b5c9",
			fixedsize = true,
			fontsize = 20,
			fontname = "Arial Narrow"
			label = synchronization,
			height = 0.7,
			//width = 2,
			shape = octagon]
		Tasks -> sync
		sched [
			fillcolor = "#6a9ab1",
			label = Scheduler]
		sync -> sched
		IRQ [
			fillcolor = "#71809b",
			fontsize = 20,
			label = "interrupts\ncore,\nCPU arch"]
		sched -> IRQ
		CPU [
			fillcolor = "#777777",
			fontcolor = white,
			fontsize = 20,
			row = chip]
		IRQ -> CPU
	}	// processing
	subgraph mem {
		node [
			color = "#51bf5b",
			height = 1
		]
		edge [color = "#51bf5b"]
		MA [
			color = "#51bf5b",
			fillcolor = "#d9e7ee",
			fixedsize = true,
			label = "memory\naccess",
			row = usr,
			height = 1,
			shape = ellipse]
		MA -> VM
		mmap [
			fillcolor = "#91b5c9",
			fixedsize = true,
			fontsize = 20,
			height = 0.8,
			label = "memory\nmapping",
			shape = octagon,
			width = 2]
		mmap -> log_mem
		log_mem -> PA
		SW [
			color = "#8383cc",
			fillcolor = "#91b5c9",
			fixedsize = true,
			label = Swap,
			height = 0.8,
			shape = octagon,
			width = 1.2]
		mmap -> SW [weight = 1]
		SW -> block [
			color = "#8383cc", weight = 1]
		PA [
			fillcolor = "#71809b",
			label = "Page\nAllocator"
		]
		PC -> PA [weight = 0 color="#51bf5b"]
		RAM [
			color = "#51bf5b",
			fillcolor = "#777777",
			fontcolor = white,
			fontsize = 20,
			label = "MMU, RAM",
			height = 1,
			row = chip]
		PA -> RAM
		memory -> memory_ [
			arrowhead = "",
			row = func]
		VM -> mmap
	}	// mem
	subgraph storage {
		node [color = "#8383cc"]
		edge [color = "#8383cc"]
		NFS;
		storage_ [
			shape = point,
			fixedsize = true,
			height = 0,
			style = invis,
			width = 0]
		storage [
			URL = "https://en.wikibooks.org/wiki/The_Linux_Kernel/Storage",
			fillcolor = white,
			fixedsize = true,
			height = 0.6,
			row = func,
			width = 2]
		storage -> storage_ [
			arrowhead = "",
			row = func]
		FS [
			fillcolor = "#d9e7ee",
			fixedsize = true,
			label = "files and\ndirectories",
			row = usr,
			shape = ellipse]
		VFS [
			fillcolor = "#b2d3e4",
			label = "Virtual\nFile System"]
		FS -> VFS
		VFS -> mmap [weight = 0]
		VFS -> NFS [weight = 0]
		logFS [
			fillcolor = "#6a9ab1",
			fontsize = 20,
			label = "logical\nfilesystems:\next3, xfs ..."]
		VFS -> logFS
		PC [
			fillcolor = "#91b5c9",
			fixedsize = true,
			fontsize = 20,
			height = 0.8,
			label = "page\ncache",
			shape = octagon,
			width = 1.2]
		VFS -> PC [weight = 0]
		block [
			fillcolor = "#71809b",
			fontsize = 20,
			label = "Block\ndevices\nand drivers"]
		logFS -> block
		SD [
			fillcolor = "#777777",
			fontcolor = white,
			fontsize = 20,
			label = "storage devices:\nSCSI, NVMe ...",
			row = chip]
		block -> SD
	}	// storge
	subgraph HI {
		node [color = "#cfbf57ff"]
		edge [
			color = "#cfbf57ff",
			weight = 10
		]
		HI_ [
			fixedsize = true,
			height = 0,
			shape = point,
			style = invis,
			width = 0]
		HI [
			URL = "https://en.wikibooks.org/wiki/The_Linux_Kernel/Human_interfaces",
			fillcolor = white,
			fixedsize = true,
			fontsize = 12,
			height = 0.6,
			label = "human\ninterface",
			row = func,
			width = 2]
		HI -> HI_ [
			arrowhead = "",
			row = func]
		char [
			fillcolor = "#d9e7ee",
			fixedsize = true,
			label = "char\ndevices",
			row = usr,
			shape = ellipse]
		input [
			fillcolor = "#b2d3e4",
			label = "input\nsubsystem"]
		char -> input
		F7 [
			fillcolor = "#6a9ab1",
			label = "HI class\ndrivers"]
		input -> F7
		HID [
			fillcolor = "#71809b",
			fontsize = 20,
			URL = "https://www.kernel.org/doc/html/latest/hid/",
			label = "HI\nperipherals\ndrivers"]
		F7 -> HID
		display [
			fillcolor = "#777777",
			fontcolor = white,
			fontsize = 19,
			label = "keyboard, mouse,\ndisplay, audio",
			row = chip]
		HID -> display
	} // HI
	subgraph functions {
		graph [rank = same]
		edge [
			style = invis,
			weight = 1
		]
		system;
		networking;
		system -> processing [weight = 1]
		storage -> networking [weight = 1]
		memory [
			color = "#51bf5b",
			URL = "https://en.wikibooks.org/wiki/The_Linux_Kernel/Memory",
			fillcolor = white,
			fixedsize = true,
			height = 0.6,
			row = func,
			width = 2]
		memory -> storage [weight = 1]
		processing -> memory [weight = 1]
		functions_ [
			fixedsize = true,
			height = 0,
			shape = point
			style = invis,
			width = 0]
		functions_ -> HI -> system [weight = 1]
		functions [
			color = gray,
			tooltip = "Columns represent main functionalities of the kernel",
			URL = "http://www.makelinux.net/ldd3/chp-1-sect-2.shtml",
			fillcolor = gray,
			fixedsize = true,
			height = 0.6,
			row = func,
			style = dashed,
			width = 1.6]
		functions -> functions_ [
			arrowhead = "",
			color = gray,
			style = "",
			weight = ""]
	}
	subgraph interfaces {
		graph [rank = same]
		SCI;
		sock;
		FS;
		proc;
		char;
		usr_ [
			fixedsize = true,
			height = 0,
			shape = point
				style = invis,
			width = 0.5]
		usr [
			fillcolor = "#d9e7eeff",
			fixedsize = true,
			label = "user space\ninterfaces",
			row = usr,
			shape = ellipse,
			style = "filled,setlinewidth(0)"]
		MA;
	}
	{
		edge [style = invis weight = 10 ]
		system_;
		SCI;
		system_ -> SCI;
		networking_;
		sock;
		networking_ -> sock;
		storage_;
		FS;
		storage_ -> FS;
		processing_;
		proc;
		processing_ -> proc;
		HI_;
		char;
		HI_ -> char;
		MA;
		memory_ [
			fixedsize = true,
			height = 0,
			shape = point,
			style = invis,
			width = 0]
		memory_ -> MA;
	}
	subgraph virtual {
		graph [rank = same]
		sysfs;
		prot_fam;
		VFS;
		Tasks;
		input;
		D0 [
			fixedsize = true,
			height = 0,
			shape = point,
			style = invis,
			width = 0]
		virt [
			fillcolor = "#b2d3e4",
			label = "virtual\nsubsystems",
			URL = "https://en.wikipedia.org/wiki/Proxy_pattern",
			tooltip = "proxy between standard user space interfaces and internal implementations",
			style = "filled,setlinewidth(0)"]
		VM [
			color = "#51bf5b",
			fillcolor = "#b2d3e4",
			label = "Virtual\nmemory"]
	}
	subgraph bridges {
		graph [rank = same]
		bridges [
			fillcolor = "#91b5c9",
			shape = octagon,
			tooltip = "bridges between uniform virtual interfaces and various implementations",
			URL = "https://en.wikipedia.org/wiki/Bridge_pattern",
			style = "filled,setlinewidth(0)"]
		DM;
		NFS;
		mmap;
		sync;
		E0 [
			fixedsize = true,
			height = 0,
			shape = point,
			style = invis,
			width = 0]
		//PC
	}
	subgraph logical {
		graph [rank = same]
		log_sys;
		log_prot;
		logFS;
		sched;
		F7;
		F0 [
			fixedsize = true,
			height = 0,
			shape = point,
			style = invis,
			width = 0]
		logical [
			fillcolor = "#6a9ab1",
			style = "filled,setlinewidth(0)"]
		log_mem [
			color = "#51bf5b",
			fillcolor = "#6a9ab1",
			label = "logical\nmemory"]
		//SW
	}
	subgraph HWI {
		graph [rank = same]
		HWI [
			fillcolor = "#71809b",
			label = "hardware\ninterfaces",
			style = "filled,setlinewidth(0)"]
		bus_drv;
		netif;
		block;
		//PA;
		IRQ;
		HID;
		G0 [
			fixedsize = true,
			height = 0,
			shape = point,
			style = invis,
			width = 0]
	}
	subgraph HW {
		graph [rank = same]
		HW [
			fillcolor = "#777777",
			fontcolor = white,
			label = "electronics,\nhardware",
			row = chip,
			style = "filled,setlinewidth(0)"]
		buses;
		net_hw;
		SD;
		CPU;
		display;
		H0 [
			fixedsize = true,
			height = 0,
			shape = point,
			style = invis,
			width = 0]
		RAM;
	}
	bottom [
		label = "© 2007-2022 Costa Shulyupin http://www.MakeLinux.net/kernel/diagram",
		URL = "http://www.MakeLinux.net/kernel/diagram",
		shape = plaintext,
		style = ""]
	CPU -> bottom [style = invis]
	layers [
		fillcolor = lightgray,
		tooltip = "Functionalities are divided to common layers. It is approximate division.",
		height = 0.1,
		style = "filled,setlinewidth(0)",
		width = 0.5]
	functions -> layers [style = invis ]
	usr -> usr_ [
		arrowhead = "",
		color = "#d9e7eeff",
		minlen = 2]
	usr -> virt [
		color = "#d9e7eeff"]
	virt -> D0 [
		arrowhead = "",
		color = "#b2d3e4",
		minlen = 2]
	virt -> bridges [
		color = "#b2d3e4"]
	bridges -> E0 [
		arrowhead = "",
		color = "#91b5c9",
		minlen = 2,
		style = "filled,setlinewidth(6)",
		weight = ""]
	bridges -> logical [
		color = "#91b5c9",
		style = "filled,setlinewidth(6)"]
	logical -> F0 [
		arrowhead = "",
		color = "#6a9ab1",
		minlen = 2,
		row = logical,
		style = "filled,setlinewidth(6)",
		weight = ""]
	logical -> HWI [
		color = "#6a9ab1",
		row = logical,
		style = "filled,setlinewidth(6)"]
	HWI -> G0 [
		arrowhead = "",
		color = "#71809b",
		minlen = 2,
		row = HWI,
		style = "filled,setlinewidth(6)",
		weight = ""]
	HWI -> HW [
		color = "#71809b",
		row = HWI,
		style = "filled,setlinewidth(6)"]
	HW -> H0 [
		arrowhead = "",
		color = "#777777",
		minlen = 2,
		row = chip,
		style = "filled,setlinewidth(6)",
		weight = ""]
	layers -> usr [
		arrowhead = "",
		color = gray,
		style = "filled,setlinewidth(1)"]
	LKD [
		fontsize = 40,
		label = "Linux kernel diagram",
		shape = plain,
		style = ""]
	LKD -> processing [style = invis]
}

```

---

Steps

1. update manuscript text
2. section order via yaml file
3. call out boxes

**
2 call out blocks/boxes to emphasize main sections of parts II (e.g., vulnerable periods, experience effects) and III (e.g., # of time points, types of stability and change). Help to add some visuals to first sections of manuscript
**

xxxxxxxxxxx
Traditional Models
  -Linear
    - Difference Scores
    - Change Scores
  -Non-Linear
    - Signed-Rank test
    - Generalized Estimating Equations

Modern Approaches
  -Linear
    - Linear Mixed Models
  -Non-Linear
    - Generalized Linear Mixed Models

SEM
  - Variable-Centered
    - Autoregressive Crosslagged Panel Model
    - Latent Change Score
    - Latent Growth Curve Models
    - Multivariate Latent Growth Curve Models
    - State-Trait Models
  - Person-Centered
    - Latent Transition Analysis
    - Latent Class Growth Analysis
  - Hybrid
    - Growth Mixture Model

Advanced SEM
  - Random-Intercept Crosslagged Panel Model
  - Latent Curve Model with Structured Residuals

Longitudinal Twin Designs/Genetic Models

Longitudinal Neuroimaging

xxxxxxxxxx\

Traditional Models
  -Linear
    - Difference Scores
    - Change Scores
  -Non-Linear
    - Signed-Rank test
    - Generalized Estimating Equations

xxxxxxxxx
    - title: "Tutorials"
      contents:
        - tutorials.qmd
        - section: "Traditional Models"
          contents:
            - 1_Tutorials_DifferenceScores.qmd
            - 2_Tutorials_ResidualizedChangeScores.qmd
            - 3_Tutorials_LinearMixedModels.qmd
        - section: "Traditional Non-linear Models"
          contents:
            - 4_Tutorials_SignedRankTest.qmd
            - 5_Tutorials_MarginalModels.qmd
            - 6_Tutorials_GeneralizedEstimatingEquations.qmd
            - 7_Tutorials_GeneralizedLinearMixedEffectsModels.qmd
        - section: "SEM Approaches"
          contents:
            - 8_Tutorials_AutoregressiveCrosslaggedPanelModels.qmd
            - 9_Tutorials_LatentChangeScoresModels.qmd
            - 10_Tutorials_LatentGrowthCurveModels.qmd
            - 11_Tutorials_MultivariateLatentGrowthCurveModels.qmd
            - 12_Tutorials_LatentTransitionAnalysis.qmd
            - 13_Tutorials_GrowthMixtureModels.qmd
            - 14_Tutorials_StateTraitModels.qmd
        - section: "Advanced SEM"
          contents:
            - 15_Tutorials_RandomInterceptCrosslaggedPanelModels.qmd
            - 16_Tutorials_LatentCurveModelsStructuredResiduals.qmd
        - section: "Longitudinal Neuroimaging"
          contents:
            - blank.qmd
        - section: "Twin Modeling & Genetics Designs"
          contents:
            - blank.qmd
        - section: "Data Wrangling"
          contents:
            - blank.qmd
    - title: "How-To"
      contents:
        - howto.qmd
        - section: "Traditional Linear Models"
          contents:
            - section: "Difference Scores"
              contents:
                - 1a_HowTos_DifferenceScores_PairedTtests.qmd
                - 1b_HowTos_DifferenceScores_SimpleRegression.qmd
            - section: "Residualized Change Scores"
              contents:
                - 2_HowTos_ResidualizedChangeScores.qmd
            - section: "Linear Mixed Models"
              contents:
                - 3a_HowTos_LinearMixedModels.qmd
                - 3b_HowTos_LinearMixedModels.qmd
        - section: "Traditional Non-linear Models"
          contents:
            - section: "Signed-Rank Test"
              contents: 
                - 4_HowTos_SignedRankTest.qmd
            - section: "Marginal Models"
              contents:
                - 5_HowTos_MarginalModels.qmd
            - section: "Generalized Estimating Equations"
              contents:
                - 6_HowTos_GeneralizedEstimatingEquations.qmd
            - section: "Generalized Linear Mixed-Effects Models"
              contents:
                - 7_HowTos_GeneralizedLinearMixedEffectsModels.qmd
        - section: "SEM Approaches"
          contents:
            - section: "Autoregressive Crosslagged Panel Models"
              contents:
                - 8_HowTos_AutoregressiveCrosslaggedPanelModels.qmd
            - section: "Latent Change Scores Models"
              contents:
                - 9_HowTos_LatentChangeScoresModels.qmd
            - section: "LatentGrowthCurveModels"
              contents:
                - 10_HowTos_LatentGrowthCurveModels.qmd
            - section: "Multivariate Latent Growth Curve Models"
              contents:
                - 11_HowTos_MultivariateLatentGrowthCurveModels.qmd
            - section: "Latent Transition Analysis"
              contents:
                - 12_HowTos_LatentTransitionAnalysis.qmd
            - section: "Growth Mixture Models"
              contents:
                - 13_HowTos_GrowthMixtureModels.qmd
            - section: "State-Trait Models"
              contents:
                - 14_HowTos_StateTraitModels.qmd
        - section: "Advanced SEM"
          contents:
            - 15_HowTos_RandomInterceptCrosslaggedPanelModels.qmd
            - 16_HowTos_LatentCurveModelsStructuredResiduals.qmd
        - section: "Longitudinal Neuroimaging"
          contents:
            - blank.qmd
        - section: "Twin Modeling & Genetics Designs"
          contents:
            - blank.qmd
        - section: "Data Wrangling"
          contents:
            - blank.qmd
        # navigation items
    - title: "Explanations"
      contents:
        - explanations.qmd
        - section: "Traditional Linear Models"
          contents:
            - 1_Explanations_DifferenceScores.qmd
            - 2_Explanations_ResidualizedChangeScores.qmd
            - 3_Explanations_LinearMixedModels.qmd
        - section: "Traditional Non-linear Models"
          contents:
            - 4_Explanations_SignedRankTest.qmd
            - 5_Explanations_MarginalModels.qmd
            - 6_Explanations_GeneralizedEstimatingEquations.qmd
            - 7_Explanations_GeneralizedLinearMixedEffectsModels.qmd
        - section: "SEM Approaches"
          contents:
            - 8_Explanations_AutoregressiveCrosslaggedPanelModels.qmd
            - 9_Explanations_LatentChangeScoresModels.qmd
            - 10_Explanations_LatentGrowthCurveModels.qmd
            - 11_Explanations_MultivariateLatentGrowthCurveModels.qmd
            - 12_Explanations_LatentTransitionAnalysis.qmd
            - 13_Explanations_GrowthMixtureModels.qmd
            - 14_Explanations_StateTraitModels.qmd
        - section: "Advanced SEM"
          contents:
            - 15_Explanations_RandomInterceptCrosslaggedPanelModels.qmd
            - 16_Explanations_LatentCurveModelsStructuredResiduals.qmd
        - section: "Longitudinal Neuroimaging"
          contents:
            - blank.qmd
        - section: "Twin Modeling & Genetics Designs"
          contents:
            - blank.qmd
        - section: "Data Wrangling"
          contents:
            - blank.qmd
        # navigation items

    - title: "Reference"
      contents:
        - reference.qmd
        - section: "List of Reference Material"
          contents:
            - ListofReferences.qmd

    - title: "Documentation"
      contents:
        - documentation.qmd

xxxxxxxxxxx

![test](./figures/Mindmap1.png)

xxx

| fruit  | price  |
|--------|--------|
| apple  | 2.05   |
| pear   | 1.37   |
| orange | 3.09   |

: Fruit prices {.striped .hover}

xxx

xxx

```{r}
library(leaflet)
leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(lng=-77.0168, lat=38.7814, popup="You are here - in Maryland!")

```

xxx

```{mermaid}

%%| fig-width: 6.5
%%| fig-height: 6.5

graph TB
A[LDA Structures] 
A -.-> |Focus on if & when an event occurs| B[Time-to-Event]
B -.-> C[Continuous-Time <br/>Survival Analysis]
B -.-> D[Discrete-Time <br/>Survival Analysis]
A -->|Focus on Change & Growth| E[Repeated Measures]
E -.-> F[Time-Series <br/> Models]
E -.-> G[Intensive <br/> Longitudinal Data]
E --> H[Longitudinal Panel Data]
H --Traditional Models--> J[Mean-level change & Marginal models]
H --Contemporary  Methods--> I[Within-person change & Random-effects models]

J --> K[ANOVA <br/> Generalized Estimating Equations <br/> ... ]
I --> L[Mixed Models <br/> Growth Curves <br/> ...]

click A href "https://www.github.com" "This is a link"
click B href "https://www.github.com" "This is a link"
click C href "https://www.github.com" "This is a link"
click D href "https://www.github.com" "This is a link"

style A fill:blue,stroke:#333,stroke-width:4px,color:#fff
style E fill:blue,stroke:#333,stroke-width:4px,color:#fff
style B fill:floralwhite,stroke:black
style C fill:floralwhite,stroke:black
style D fill:floralwhite,stroke:black
style F fill:floralwhite,stroke:black
style G fill:floralwhite,stroke:black

style H fill:blue,stroke:#333,stroke-width:4px,color:#fff
style J fill:blue,stroke:#333,stroke-width:2px,color:#fff
style I fill:blue,stroke:#333,stroke-width:2px,color:#fff
style K fill:blue,stroke:#333,stroke-width:2px,color:#fff,text-align:center
style L fill:blue,stroke:#333,stroke-width:2px,color:#fff,text-align:center

```

![FigmaLandingTest2](./figures/Figlanding5.svg) (remove marginal models from "examples" address (vs random-effects) in the manuscript)

-   section: "Traditional Linear Models" contents: - 1_Tutorials_DifferenceScores.qmd - 2_Tutorials_ResidualizedChangeScores.qmd - section: "Traditional Non-linear Models" contents: - 4_Tutorials_SignedRankTest.qmd - 6_Tutorials_GeneralizedEstimatingEquations.qmd (population average effects) - section: "Modern GLM Extensions(?)" - 7_Tutorials_LinearMixedModels.qmd\
    -   8_Tutorials_GeneralizedLinearMixedEffectsModels.qmd - section: "SEM Approaches" contents: - 8_Tutorials_AutoregressiveCrosslaggedPanelModels.qmd - 9_Tutorials_LatentChangeScoresModels.qmd - 10_Tutorials_LatentGrowthCurveModels.qmd - 11_Tutorials_MultivariateLatentGrowthCurveModels.qmd - 12_Tutorials_LatentTransitionAnalysis.qmd - 13_Tutorials_GrowthMixtureModels.qmd - 14_Tutorials_StateTraitModels.qmd - section: "Advanced SEM" contents: - 15_Tutorials_RandomInterceptCrosslaggedPanelModels.qmd - 16_Tutorials_LatentCurveModelsStructuredResiduals.qmd

GEE, on the other hand, is fitting a marginal model. These model population-averages. You're modeling the expectation conditional only on your fixed design matrix. GLMM is fitting a mixed-effects model; Thus the the estimates are subject-specific. What differs between GEE and GLMM is the target of inference: population-average or subject-specific.

This is in contrast to mixed effect models as explained above which condition on both the fixed design matrix and the random effects. So with the marginal model above you're saying, "forget about the difference among classrooms, I just want the population (school-wise) rate of failure and its association with gender." You fit the model and get an odds ratio that is the population-averaged odds ratio of failure associated with gender.

Molenberghs, Verbeke 2005 has an entire chapter on marginal vs. random effects models.

xxxxxx

|                                           |                         |                          |                                      |                         |
|:-----------------------------|:----------------------:|:---------------------:|:---------------------------:|:-----------------:|
| Models                                    |     Form of Change      | Minimum \# of Timepoints | Variable-Centered vs Person-Centered | Time-varying Covariates |
| Difference Scores                         |       mean-level        |            2+            |                  \-                  |          none           |
| Change Scores                             |       rank-order        |            2+            |                  \-                  |          none           |
| Signed-Rank Test                          |       rank-order        |            2+            |                  \-                  |          none           |
| Generalized Estimating Equations          |       rank-order        |            2+            |                  \-                  |         allowed         |
| Linear Mixed Models                       |   within and between    |            3+            |          Variable-Centered           |         allowed         |
| Generalized Linear Mixed-Effects Models   |   within and between    |            3+            |                                      |         allowed         |
| Autoregressive Crosslagged Panel Models   |       rank-order        |            3+            |          Variable-Centered           |         allowed         |
| Latent Change Scores Models               |       rank-order        |            2+            |          Variable-Centered           |         allowed         |
| Latent Growth Curve Models                |   within and between    |            3+            |          Variable-Centered           |         allowed         |
| Multivariate Latent Growth Curve Model    |   within and between    |            3+            |          Variable-Centered           |         allowed         |
| Latent Transition Analysis                |   within and between    |            3+            |                Mixed                 |         allowed         |
| Latent Class Growth Analysis              |   within and between    |            3+            |           Person-Centered            |         allowed         |
| Growth Mixture Models                     |   within and between    |            3+            |                Mixed                 |         allowed         |
| State-Trait Models                        | continuous and discrete |            3+            |          Variable-Centered           |         allowed         |
| Random-Intercept Crosslagged Panel Models |   within and between    |            4+            |          Variable-Centered           |         allowed         |
| Latent Curve Models Structured Residuals  |   within and between    |            4+            |          Variable-Centered           |         allowed         |
|                                           |                         |                          |                                      |                         |
: Models for the analysis of longitudinal data {.striped .hover .responsive-xl}

xxxxxx

2.  Models for Longitudinal Data Since both the GLMM and WGEE are extensions of the GLM, we start with a brief overview of the latter.

2.1 Generalized Linear Models (GLM)

LMM vs LGCM? For modeling growth, researchers typically employ some type of random effects model such that a mean growth trajectory is estimated for all observations in the data but a unique growth curve is estimated for each individual in the data as well (Curran & Bauer, 2011). These types of models are generally referred to as subject-specific models (Zeger, Liang, & Albert, 1988), which are more commonly known in psychology as growth curve models. Growth models have many aliases but can be broadly grouped into two different classes of methods: the latent-curve (LC) approach, which treats the repeated measures as multivariate (also known as the "wide" data format) and tends to be fit with general structural equation modeling (SEM) software (Meredith & Tisak, 1990; Tucker, 1958; Willett & Sayer, 1994), and the mixed-effect (ME) approach, which treats the repeated measures as univariate (also known as the "long" data format) and is generally fitted with regression software (Bryk & Raudenbush, 1987; Laird & Ware, 1982; Rao 1965).Footnote1

Over the past 20 years, methodological research has shown that the LC and ME approaches are actually nuanced twists on the same idea; they have been shown to converge, and to be mathematically equivalent in many cases (e.g., Bauer, 2003; Curran, 2003; Ledermann & Kenny, 2017; Mehta & Neale, 2005).

::: page-layout-custom
![FigmaLandingTest2](./figures/Figlanding2.svg)
:::

xxxxx It is important to match a theory of change with design and statistical model to assess change. (p. 4) Models with two time points cannot provide estimates of within-person change. "Stability" over time can reflect many things, only some of which mean a lack of change. (p. 14) It is helpful to describe not only average growth (e.g. intercept and slope means), but variances in the growth parameters. (p. 25)

Models that focus exclusively on mean-level change sometimes called Marginal Models Those that emphasize individual change often do so through inclusion of Random Effects often called "growth models"

xxxxx

```{mermaid}
mindmap
  root((mindmap))
    Origins
      Long history
      ::icon(fa fa-book)
      Popularisation
        British popular psychology author Tony Buzan
    Research
      On effectiveness<br/>and features
      On Automatic creation
        Uses
            Creative techniques
            Strategic planning
            Argument mapping
    Tools
      Pen and paper
      Mermaid

```

xxxxx

Longitudinal Psychometrics

xxxxxx

::: {style="color: red; border: solid black; background-color: grey;"}
This is the [middle section]{style="background-color: yellow; color: blue;"}
:::

::: grid
::: g-col-2
:::

::: g-col-8
::: {style="color: red; border: solid black; background-color: green;"}
Testing [Sample Test!]{style="background-color: yellow; color: blue;"}
:::
:::

::: g-col-2
:::
:::

The primary goal of the tutorial section is to help users learn and understand a specific topic or method by providing an introductory and hands-on, step-by-step guide. This section aims to create engaging, accessible, and effective entry learning experiences that enable users to develop their skills and understanding of the subject matter.

<br>

::: grid
::: g-col-2
:::

::: g-col-8
![](./figures/lights.png)
:::

::: g-col-2
:::
:::

::: {style="text-align: justify; text-align: center; line-height: normal; color: white; border: solid black; background-color: darkgoldenrod; padding: 70px; font-family: Georgia, serif;"}
This is the temporary landing page for the Longitudinal Analysis Project. <br> The site is currently under construction and used for testing purposes only. <br> Contents may be inaccurate.
:::

<br>

::: grid
::: g-col-2
:::

::: g-col-8
![](./figures/lights.png)
:::

::: g-col-2
:::
:::

</div>