### 2023-06-15 _data_wrangling.qmd
revised text

# **Data Wrangling** {.unnumbered}

```{r}
#| message: FALSE
#| warning: FALSE
#| echo: FALSE

#Clear Environment
#rm(list=ls())
```

<style>
div.blue pre { background-color:ghostwhite; }
div.blue pre.r { background-color:seashell; }

.nav-tabs > li.active:nth-child(1) > a {
  background-color: ##964B00;
}

.nav-tabs > li > a {
  background-color: ##FFFF00;
  color: blue;
}

main-container {
   max-width: unset;
}
.descriptives {
   overflow: auto; /*do not use `resize`*/
}

.tableby {
   overflow: auto;
}

.descriptives {
   overflow: auto;
   resize: both;
}

table, th, td {
  padding-top: 0px !important;
  padding-bottom: 0px !important;
  vertical-align: middle;
}
</style>

::: {.column-body-outset}
## Preliminary Setup
::: {.panel-tabset}
### Load Libraries {.tabset .tabset-fade .tabset-pills}
example-text before code block
<div class = "blue">
```{r pckg-install}
#| echo: TRUE
#| messages: FALSE
#| warning: FALSE

## Install necessary packages (if not already installed)
if (!("tidyverse" %in% installed.packages())) install.packages("tidyverse")
if (!("arrow" %in% installed.packages())) install.packages("arrow")

```
</div>
example-text after code block

### Load Packages
<div class = "blue">
```{r pckg-load}
#| echo: TRUE
#| messages: FALSE
#| warning: FALSE
#| class.source: fold-hide

## Load packages
library(tidyverse)
library(arrow)
library(tippy)
```
</div>

### Config Options
<div class = "blue">
```{r config}
#| echo: TRUE
#| messages: FALSE
#| warning: FALSE

knitr::opts_chunk$set(echo = T, message=F, warning=F, error=F, 
                      comment=NA, cache=T, code_folding=T,
                      R.options=list(width=220), fig.align='center', 
                      out.width='75%', fig.asp=.75)
```
</div>

```{r}
##| include: FALSE
##| evaluate: FALSE
#| message: FALSE
#| warning: FALSE
#| echo: FALSE

#df_ndaFile <- readRDS("/Users/shawes/Library/CloudStorage/GoogleDrive-swh004@gmail.com/My Drive/bio_Main/_data/nda4.0.rds")

#df_test <- readRDS("path_to_ndaFile/nda4.0.rds")

#Create temporary dataframe
#df_temp <- (df_ndaFile) %>%
#   select(subjectid, eventname, site_id_l, rel_family_id, sex_at_birth, 
#          demo_brthdat_p, screentime_13,screentime_wkdy_3, screentime_wknd_9,
#          cbcl_scr_syn_aggressive_r, cbcl_scr_syn_external_r)

# remove large dataframe object

#write_dataset(df_temp, "./df_temp", format = "csv")
df_temp<- read_csv("./df_temp.csv")

df_temp <- df_temp %>% rename("ids" = "subjectid",
                    "event" = "eventname",
                    "site_id" = "site_id_l",
                    "fam_id"="rel_family_id",
                    "sex"="sex_at_birth",
                    "age"="demo_brthdat_p",
                    "mature_vg" = "screentime_13",
                    "vg_wkdy" = "screentime_wkdy_3",
                    "vg_wknd" = "screentime_wknd_9",
                    "cbcl_agg"="cbcl_scr_syn_aggressive_r",
                    "cbcl_extern"="cbcl_scr_syn_external_r"
                    )

df_temp$ids <- as.factor(df_temp$ids)
df_temp$event <- as.factor(df_temp$event)
df_temp$site_id <- as.factor(df_temp$site_id)
df_temp$fam_id <- as.factor(df_temp$fam_id)
df_temp$sex <- as.factor(df_temp$sex)
df_temp$age <- as.numeric(df_temp$age)
df_temp$mature_vg <- as.numeric(df_temp$mature_vg)
df_temp$vg_wkdy <- as.numeric(df_temp$vg_wkdy)
df_temp$vg_wknd <- as.numeric(df_temp$vg_wknd)
df_temp$cbcl_agg <- as.numeric(df_temp$cbcl_agg)
df_temp$cbcl_extern <- as.numeric(df_temp$cbcl_extern)

## recode value of each level for factors
df_temp$sex<-recode(df_temp$sex, 'F' = 'Female', 'M' = 'Male')

df_temp$event<-recode(df_temp$event, 'baseline_year_1_arm_1' = 'Baseline',
                      '6_month_follow_up_arm_1' = '6_month', 
                      '1_year_follow_up_y_arm_1'= 'Year_1', 
                      '18_month_follow_up_arm_1' = '18_month',
                      '2_year_follow_up_y_arm_1'='Year_2', 
                      '30_month_follow_up_arm_1' = '30_month',
                      '3_year_follow_up_y_arm_1'='Year_3',
                      '42_month_follow_up_arm_1' = '42_month'
                      )

## relevel factor categories
df_temp$event <- fct_relevel(df_temp$event, "Baseline", "Year_1", "Year_2")

df_temp<-df_temp[!(df_temp$event=="6_month" | df_temp$event=="18_month" |
                     df_temp$event=="30_month" | df_temp$event=="42_month" | 
                     df_temp$event=="Year_3"),]

#drop unused factor levels
df_temp$event <- droplevels(df_temp$event)
levels(df_temp$event)

## Compute total video gaming hours variable
df_temp$vg_total <- (df_temp$vg_wkdy*5)+(df_temp$vg_wknd*2)
df_temp$vg_total
## Compute video gaming grouping variable
df_temp$vg_group <- ifelse(df_temp$vg_total>=7, "VG", NA)
df_temp$vg_group <- ifelse(df_temp$vg_total>0 & df_temp$vg_total<7, "NVG", df_temp$vg_group)
## set video gaming grouping variable as factor variable
df_temp$vg_group <- as.factor(df_temp$vg_group)

##############
#creating 3 separate dataframes for each "event" period (Baseline, Year_1, Year_2)
# due to uncertainty in best procedure for converting original "long-format" of the 
# dataset to wide format (as needed for SEM analyses) when filtering participants at each
# "event" based on their video game playing hours only at "event"== "baseline"

## create T0 dataframe, filtering by chosen criterion
df_t0 <- df %>%
  filter(df$event == "Baseline" & df$vg_total >= 7)
summary(df_t0)

## create T1 dataframe, filtering by chosen criterion
df_t1 <- df %>%
  filter(df$event == "Year_1")
df_t1 <- (df_t1) %>%
  select(ids, cbcl_agg, cbcl_extern)
summary(df_t1)

## create T2 dataframe, filtering by chosen criterion
df_t2 <- df %>%
  filter(df$event == "Year_2")
df_t2 <- (df_t2) %>%
  select(ids, cbcl_agg, cbcl_extern)
summary(df_t2)

df_merge <- merge(df_t0, df_t1, by='ids')
df_merge <- merge(df_merge, df_t2, by='ids')
df_merge <- df_merge %>% rename("cbcl_agg_T2" = "cbcl_agg",
                                "cbcl_agg_T0" = "cbcl_agg.x",
                                "cbcl_agg_T1" = "cbcl_agg.y",
                                "cbcl_extern_T2" = "cbcl_extern",
                                "cbcl_extern_T0" = "cbcl_extern.x",
                                "cbcl_extern_T1" = "cbcl_extern.y"
                                )
summary(df_merge)
print(df_merge)
#df_xxxxx <- df_xxxxx %>%
#filter(mature_vg > 1)

******************************

df_wide <- (df_merge) %>%
  select(ids, event, site_id, fam_id, sex, age, mature_vg, vg_total, vg_group,
         cbcl_agg_T0, cbcl_agg_T1, cbcl_agg_T2,
         cbcl_extern_T0, cbcl_extern_T1, cbcl_extern_T2)

###########

df_sub <- df %>% group_by(event) %>% sample_n(500)
print (df_sub)
###########

write_dataset(df, "./df", format = "csv")

rm(df_ndaFile, df_temp)
```

```{r}
#| message: FALSE
#| warning: FALSE
#| echo: FALSE

#create a new 'long-to-wide' tibble from original tibble

#df_wide <- pivot_wider(df, id_cols = ids,
#                       names_from = event, 
#                       values_from = cbcl_agg,
#                       )
#head(df_wide)

#mature_vg, vg_total, vg_group, cu_traits, DBD, DBDcu, 
#cbcl_rulbrk, cbcl_agg, cbcl_extern, cbcl_opp, cbcl_cd

#as_tibble(df_wide)
#tibble(df_wide)

#write_dataset(df_wide, "./", format = "feather")
#write_dataset(df_wide, "./", format = "csv")
```

```{r}
#| class.source: 'fold-hide'
#| message: FALSE
#| warning: FALSE
#| echo: FALSE

df_long <- df_wide %>%
  pivot_longer(
    cols = starts_with("cbcl_agg"),
    names_to = "time",
    names_prefix = "cbcl_agg_",
    values_to = "cbcl_agg"
  )

df_long <- (df_long) %>%
  select(ids, time, site_id, fam_id, sex, age, mature_vg, vg_total, vg_group, 
         cbcl_agg)

 df_long <- df_long %>% rename("event" = "time")
 df_long$event <- as.factor(df_long$event)
  df_long

df_long_sub <- df_long %>% sample_n(500)
print (df_long_sub)

#create a new 'wide-to-long' tibble from original tibble
# longer <- df_wide %>% 
#     pivot_longer(cols = -ids, 
#                  names_to = c(".value", "event2"), 
#                  names_pattern = "([^\\.]*)\\.*(\\{4})")
# 
# longer<-pivot_longer(df_wide, cols=-1, names_pattern = "(.*)(..)$", names_to = c("limit", "name")) %>% 
#      mutate(limit=ifelse(limit=="", "value", limit))
# 
# answer <-pivot_wider(longer, id_cols = c(event, name), names_from = limit, values_from = value, names_repair = "check_unique")

# df_long <- df_wide %>%
#   pivot_longer(
#     cols = starts_with("cu_traits"),
#     names_to = "time",
#     names_prefix = "cu_traits_",
#     values_to = "cu_traits"
#   )
# 
# df_long <- (df_long) %>%
#   select(ids, time, site_id, fam_id, sex, age, mature_vg, vg_total, vg_group,
#          DBD, DBDcu, cu_traits) 
# 
#  df_long <- df_long %>% rename("event" = "time")
#  df_long$event <- as.factor(df_long$event)
# 
# 
# df_long <- df_long
write_dataset(df_long, "./df_long", format = "csv")
write_dataset(df_long_sub, "./df_long_sub", format = "csv")
```